---
title: "Data Integration Script"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
In this rmd, I take SMFS data developed in the SMFSDataTableMerging and SMFS_Wrangling rmds and integrate with SBOTS data wrangled and checked by Lewis. 
The goal of this rmd is to create a combined dataset that can be used for incubator projects and beyond. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Merging the SMFS and SBOT data.

## Reading in the data and loading libraries

```{r}
library(readxl)
library(tidyverse)
library(readr)

SBOTS2019 <- read_excel("Data/SBOTS2019.xlsx")
SMFS <- read_csv("Data/SMFS_Integrate.csv")
```

Wrangling

We want our end table to have distinct rows for SMFS and SBOT fishes but the catch and wq info goes into the same column.

# 1. Need to add column for "study" that specifies which study the data came from.

```{r}
#Need to add a column called Study to make it easy to parse later after merge.
SMFS<-mutate(SMFS, Study = "SMFS")
SBOTS2019<-mutate(SBOTS2019, Study = "SBOTS")
```

# 2. Matching columns...

Left is SMFS, right is SBOTS. All water quality taken by SMFS is done for the surface of the water column so all water quality from SBOTS will start with "es" indicating it was taken at the end of the trawl at the surface of the water, matching SMFS methodology.

|                     |                                        |           |                           |                                                                                                                                                                                                                                    |
|---------------|---------------|---------------|---------------|---------------|
| **SMFS**            | **SMFS_Example**                       | **SBOTS** | **SBOT_Example**          | **Notes**                                                                                                                                                                                                                          |
| TrawlRowID          | {00002451-7B5A-47A8-8D9F-486D054DC257} | id        | 3654                      | Unique identification number assigned  to each tow when entered into the database. Not all data entered in order of date, so id \# may be out of order with date. Multiple rows have same id for each species/length combination.  |
| WaterTemperature    | 11.8                                   | estemp    | 12.46                     | Water temperature in degrees C.                                                                                                                                                                                                    |
| Secchi              | 49                                     | esecchi   | 36                        | Measure of water clarity in cm.                                                                                                                                                                                                    |
| Salinity            | 4.9                                    | essal     | 10.63                     | Salinity measured in ppt (but actually practical units?).                                                                                                                                                                          |
| TowDuration         | 5                                      | dur       | 10                        | Number of minutes of tow (sampling effort).                                                                                                                                                                                        |
| DO                  | 8.41                                   | esdoc     | 10.05                     | Dissolved oxygen measured in mg/L.                                                                                                                                                                                                 |
| Count               | 12                                     | count     | 16                        | Number of organism counted in sample.                                                                                                                                                                                              |
| PctSaturation       | 80.2                                   | esdop     | 92.3                      | Dissolved oxygen measured in %.                                                                                                                                                                                                    |
| SpecificConductance | 8670                                   | escond    | 17935                     | Specific conductivity (μS).                                                                                                                                                                                                        |
| StationCode         | MZ2                                    | stn       | ALV 1                     | Name of location of sample.                                                                                                                                                                                                        |
| MethodCode          | OTR                                    | method    | LOT                       | Methodology used to collect sample.                                                                                                                                                                                                |
| SampleDate          | 12/4/2013                              | towdate   | 1/12/2019                 | Date of sample.                                                                                                                                                                                                                    |
| TideCode            | flood                                  | tidet     | Ebb                       | Type of tide.                                                                                                                                                                                                                      |
| TideHeight          | NA                                     | tideh     | Ex: 5.3 (Unsure of units) | Height of tide at location for the date and time. (Will need to be calculated for SMFS).                                                                                                                                           |
| Year                | 2012                                   | year      | 2019                      | Year of sample.                                                                                                                                                                                                                    |
| WaterYear           | 2011                                   | wyr       | 2018                      | Water year of sample.                                                                                                                                                                                                              |
| StandardLength      | 63                                     | sl        |                           | Not currently in SBOT table but may be included later.                                                                                                                                                                             |
| Month               | 12                                     | month     | 1                         | Month of sample.                                                                                                                                                                                                                   |
| CommonName          | Arrow Goby                             | comname   | Mississippi Silverside    | Common name of species.                                                                                                                                                                                                            |
| Family              | Gobiidae                               | family    | Atherinopsidae            | The family of the species.                                                                                                                                                                                                         |
| Genus               | Clevelandia                            | genus     | Menidia                   | Genus of the species.                                                                                                                                                                                                              |
| gensp               | Clevelandia ios                        | gensp     | Menidia audens            | Genus species.                                                                                                                                                                                                                     |
| Native              | TRUE or FALSE                          | natinv    | Invasive                  | Is species native or not/invasive?                                                                                                                                                                                                 |
| Study               | SMFS                                   | Study     | SBOTS                     | Which study collected this data?                                                                                                                                                                                                   |

```{r}
#Checking column types match up
glimpse(SMFS)
glimpse(SBOTS2019)
#Fixing column types so they work together.
SBOTS2019$estemp<-as.numeric(SBOTS2019$estemp)
SBOTS2019$esecchi<-as.numeric(SBOTS2019$esecchi)
SBOTS2019$essal<-as.numeric(SBOTS2019$essal)
SBOTS2019$esdoc<-as.numeric(SBOTS2019$esdoc)
SBOTS2019$esdop<-as.numeric(SBOTS2019$esdop)
SBOTS2019$escond<-as.numeric(SBOTS2019$escond)
SMFS$Native<-as.character(SMFS$Native)
glimpse(SBOTS2019)
```

# 3. Join

```{r}
#Joining the two datasets.
Integrated<-full_join(SMFS, SBOTS2019, join_by(WaterTemperature==estemp, Secchi==esecchi, CPUE==cpue, Salinity==essal, TowDuration==dur, DO==esdoc, Count==count, PctSaturation==esdop, SpecificConductance==escond, StationCode==stn, MethodCode==method, SampleDate==towdate, TideCode==tidet, Year==year, WaterYear==wyr, Month==month, CommonName==comname, Family==family, Genus==genus, Species==species, Native==natinv, gensp==gensp, Study==Study)) #join_by allows for easy matching of columns with different names. Retains first name.

Reduced<-Integrated %>% select(CommonName, Family, Genus, Species, gensp, Native, StationCode, Study, MethodCode, TowDuration, Count, CPUE, WaterTemperature, Secchi, Salinity, DO, PctSaturation, SpecificConductance, TideCode, SampleDate, Year, WaterYear, Month) #Selecting only the columns of interest. Standard length will be included later.

#Save New DF
#write.csv(Reduced, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/IntegratedData.csv")
```
