---
title: "Modeling Protocol"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Constructed so I can search "LongfinSmelt"and replace with Gensp name and have the code run.

### Loading libraries and data

```{r}
library(readr)
library(dismo)
library(caret)
library(here)
library(cowplot)
library(Hmisc)
library(tidyverse)

PresAbs<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_PresAbs.csv")
```

# PresenceAbsence

## Step 1: Data visualization, Quality Check

Before modeling, we will ensure the data is up to snuff.

First things first, we need to get the data for the LongfinSmelt of interest we want to model.

```{r}
LongfinSmeltData<-PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys")
LongfinSmeltData<-as.data.frame(LongfinSmeltData)
summary(LongfinSmeltData$TowDuration)
```

Now lets check the dataset.

```{r}
glimpse(LongfinSmeltData)
summary(LongfinSmeltData)
```

Let's make a couple of plots for making predictions.

```{r}
# I want to show presences vs. all samples.

Plot_Water<-LongfinSmeltData %>% ggplot(aes(y=PresAbs, x=WaterTemperature, color = TowDuration)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw() + ylab("") + xlab("Water Temperature (C)") #+ facet_wrap(~Year)

Plot_DO<-LongfinSmeltData %>% ggplot(aes(y=PresAbs, x=DO, color = TowDuration)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("")+ xlab("Dissolved Oxygen (mg/L)") #+ facet_wrap(~Year)

Plot_Secchi<-LongfinSmeltData %>% ggplot(aes(y=PresAbs, x=Secchi, color = TowDuration)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Secchi (cm)") #+ facet_wrap(~Year)

Plot_Salinity<-LongfinSmeltData %>% ggplot(aes(y=PresAb, x=Salinity, color = TowDuration)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Salinity") #+ facet_wrap(~Year)

Plots<-plot_grid(Plot_Water, Plot_DO, Plot_Secchi, Plot_Salinity, ncol = 2, labels = "AUTO")

# now add the title
title <- ggdraw() + draw_label("Longfin smelt presence/absence by water quality variables for SMFS 2011-2023", fontface = 'bold', x = 0, hjust = 0) +  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7))

plot_grid(title, Plots, ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1))

SummTable <- LongfinSmeltData %>% 
  group_by(PresAbs) %>% #break down yearly summaries by including Year as a grouping variable. 
  dplyr::summarize(
    Mean_WT = mean(WaterTemperature),
    SD_WT = sd(WaterTemperature),
    MinWT=min(WaterTemperature),
    MaxWT=max(WaterTemperature),
    Mean_DO = mean(DO),
    SD_DO = sd(DO),
    MinDO=min(DO),
    MaxDO=max(DO),
    Mean_Secchi = mean(Secchi),
    SD_Secchi = sd(Secchi),
    MinSecchi=min(Secchi),
    MaxSecchi=max(Secchi),
    Mean_Salinity = mean(Salinity),
    SD_Salinity = sd(Salinity),
    MinSal=min(Salinity),
    MaxSal=max(Salinity)
  ) 

print(SummTable)
write.csv(SummTable, here("data", "LongfinSmelt_SummTable.csv"))

#Seasonality
LongfinSmeltData$Year<-as.factor(LongfinSmeltData$Year)

LongfinSmeltData$Month<-as.factor(LongfinSmeltData$Month)

LongfinSmeltData$StationCode<-as.factor(LongfinSmeltData$StationCode)

LongfinSmeltData %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n))

LongfinSmeltData %>% group_by(Month) %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=Month, y=TotalPresence)) +geom_col()+geom_linerange(aes(ymin=TotalPresence-sd_binary, ymax=TotalPresence+sd_binary))+theme_bw(base_size = 26)+ylab("Trawls that catch longfin smelt") +xlab("Month (summed 2011-2023)") #+ theme(axis.text.x = element_text(angle = 45, hjust=1))
LongfinSmeltData %>% group_by(StationCode) %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=StationCode, y=PercentageOfPresence)) +geom_col()+theme_bw(base_size = 26)+ylab("% of trawls that catch longfin smelt") +xlab("Station Code") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```


Plot analysis: Average water temperature when LFS absent is 16.37 C (standard deviation is 4.64), present is 17.36 C (standard deviation 3.14). Average dissolved oxygen when LFS absent is 6.95 mg/L (sd 1.64), present is 7.39 (sd 1.18). Average secchi when LFS absent is 30.5 cm (sd 11.55), present is 30.56 cm (sd 10.18). Average salinity when LFS absent is 4.66 psu (sd 3.4) and present is 5.4 (sd 3.24).

Range for presences: 7.6 to 23.8 C; 4.2 to 10.10 mg/L; 11 to 73 cm; 0.2 to 14.2 psu.

Range for absences (other samples): 4.6 to 27 C; 1 to 13.74 mg/L; 7 to 99; 0 to 15 psu.

Most restricted presence compared to absence is dissolved oxygen. High secchi depth (clear water) also lacks presence above 75 cm.

Do the most influential?

## Step 2: Use gbm() from dismo to run model and create outputs

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

```{r}
 set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
system.time( brt1 <- dismo::gbm.step(data=LongfinSmeltData, 
                         gbm.x= c(21, 22, 23, 25), # environmental variables. 29 is tow duration if you want to add that back in.
                         gbm.y= 17, # response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = 3, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = 0.005,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = 0.6 # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brt1)

names(brt1) #tells us the components of output

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)

 # reponse curves
par(mfrow=c(3,2))

dismo::gbm.plot(brt1, n.plots = 5, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,3), common.scale = T, y.label = "Fitted Value")

gbm::plot.gbm(brt1, n.plots=5, return.grid=TRUE)

dismo::gbm.plot.fits(brt1)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brt1)
find.int
brt1$self.statistics$correlation[[1]]
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brt1,1, 2) #3D PLOT of Water Temp and Salinity
gbm.perspec(brt1,3, 2) #salinity and DO
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brt1, here("data", "LongfinSmelt_brt_gbm_step.rds"))

```

-   tree.complexity = 5, learning.rate = 0.001, bag.fraction = 0.6 (Nima's preference) fit final gbm model with a fixed number of 5800 trees for Binary mean total deviance = 0.467 mean residual deviance = 0.304

estimated cv deviance = 0.377 ; se = 0.008

training data correlation = 0.546 cv correlation = 0.359 ; se = 0.024

training data AUC score = 0.921 cv AUC score = 0.828 ; se = 0.014

elapsed time - 0.03 minutes user system elapsed 92.58 0.03 105.67 - Increase bag fraction to the recommendation from Elith et al of 0.75: fit final gbm model with a fixed number of 6150 trees for Binary

mean total deviance = 0.467 mean residual deviance = 0.302

estimated cv deviance = 0.373 ; se = 0.01

training data correlation = 0.552 cv correlation = 0.368 ; se = 0.026

training data AUC score = 0.923 cv AUC score = 0.831 ; se = 0.012

elapsed time - 0.03 minutes user system elapsed 88.67 0.14 110.60

-   learning rate 0.05 attempted but too big. Has to be smaller to reach minimum

-   learning rate 0.005: fit final gbm model with a fixed number of 1500 trees for Binary

mean total deviance = 0.467 mean residual deviance = 0.291

estimated cv deviance = 0.377 ; se = 0.014

training data correlation = 0.574 cv correlation = 0.36 ; se = 0.04

training data AUC score = 0.93 cv AUC score = 0.835 ; se = 0.015

elapsed time - 0.46 minutes user system elapsed 20.75 0.05 27.56

-   changed bag fraction to 0.6 fitting final gbm model with a fixed number of 1600 trees for Binary

mean total deviance = 0.467 mean residual deviance = 0.285

estimated cv deviance = 0.373 ; se = 0.01

training data correlation = 0.582 cv correlation = 0.383 ; se = 0.019

training data AUC score = 0.934 cv AUC score = 0.829 ; se = 0.014

Firstly, the things you can see: The R console will show some results (see the Word Working Guide of the Elith tutorial for an example). It reports a brief model summary & all the values are also retained in the model object, so they will be permanently kept (as long as you save the R workspace before quitting).

There will also be a graph.. This model was built with the default 10-fold cross-validation (CV). The solid black curve is the mean, and the dotted curves +- 1 standard error, for the changes in predictive deviance (ie as measured on the excluded folds of the CV). The red line shows the minimum of the mean, and the green line the number of trees at which that occurs. The final model that is returned in the model object is built on the full data set, using the number of trees identified as optimal.

For the final model (lr = 0.005, tc = 5, bag fraction = 0.6)

Interaction report:

find.int \$rank.list

| var1.index | var1.names | var2.index | var2.names       | int.size |
|------------|------------|------------|------------------|----------|
| 2          | Salinity   | 1          | WaterTemperature | 169.93   |
| 3          | DO         | 2          | Salinity         | 111.24   |

\

|             | Water Temp | Salinity | DO     | Secchi | TowDuration |
|-------------|------------|----------|--------|--------|-------------|
| Water Temp  | 0          | 169.93   | 44.64  | 23.68  | 0.63         |
| Salinity    | 0          | 0        | 111.24 | 30.97  | 7.6         |
| DO          | 0          | 0        | 0      | 14.93  | 13.83       |
| Secchi      | 0          | 0        | 0      | 0      | 3.61        |
| TowDuration | 0          | 0        | 0      | 0      | 0           |

## Validation

Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = 0.005, k_folds = 10, tree.complexity = 3, bag.fraction = 0.6){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltData,
                        gbm.x = c(21, 22, 23, 25),
                        gbm.y=17, learning.rate = 0.005,
                        bag.fraction = 0.6, tree.complexity = 3,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = 0.005, tree.complexity = 3, bag.fraction = 0.6){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltData[which(LongfinSmeltData$Binary == 1),],
                        abs = LongfinSmeltData[which(LongfinSmeltData$Binary == 0),],
                        gbm.x = c(21, 22, 23, 25),
                        gbm.y=17, learning.rate = 0.005,
                        bag.fraction = 0.6, tree.complexity = 3)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval <- list(brt = brt1, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval
#now save it

saveRDS(eval, here("data","LongfinSmelt_brt_SDM_eval_gbm_step.rds"))
```

-   learning.rate = 0.005, bag.fraction = 0.75, tree.complexity = 5: 
\$brt

A gradient boosted model with bernoulli loss function. 1700 iterations were performed. There were 5 predictors of which 5 had non-zero influence.

\$brt_k 
k Deviance AUC TSS 
1 39.28567 0.660442 0.4690653 
2 41.62835 0.7488567 0.4283537 
3 40.37159 0.8390805 0.5609979 
4 40.14176 0.7623641 0.4362772 
5 33.89985 0.8547767 0.6200828 
6 36.50092 0.8283688 0.5440729 
7 37.60592 0.8617647 0.6368421 
8 34.83711 0.9014946 0.7039402 
9 35.42474 0.8515509 0.5520791 
10 36.62672 0.8634146 0.6380081

\$brt_loo k Deviance AUC TSS n_pres 
2013 42.18586 0.7363281 0.4687500 182 
2021 39.57273 0.8402778 0.6388889 178 
2017 46.13577 0.6388889 0.5000000 208 
2014 60.69164 0.7396694 0.4545455 192 
2022 36.61205 0.8592593 0.6000000 169 
2015 46.60470 0.7142857 0.4285714 207 
2011 42.22235 0.5739645 0.2307692 201 
2020 41.28407 0.9318182 0.7727273 192 
2012 52.62661 0.9387755 0.8571429 207 
2023 47.37607 0.9218750 0.7500000 206 
2018 45.84699 0.8055556 0.5833333 202 
2019 45.00243 0.7500000 0.5000000 212 
2016 46.18059 0.7500000 0.5000000 212

-   bag fraction = 0.6 A gradient boosted model with bernoulli loss function. 1600 iterations were performed. There were 5 predictors of which 5 had non-zero influence. 
\$brt_k 
k Deviance AUC TSS 
1 1 33.73731 0.8855978 0.7383152 
2 2 39.09204 0.8116568 0.5367872 
3 3 34.86225 0.8910326 0.6425272 
4 4 38.39483 0.7958204 0.4708978 
5 5 37.40607 0.8377709 0.5851393 
6 6 34.76156 0.8281326 0.5079826 
7 7 36.93593 0.7928128 0.4803313 
8 8 42.00954 0.8098272 0.5533843 
9 9 35.63471 0.8245614 0.5630252 
10 10 38.26671 0.8130731 0.5434783

\$brt_loo k Deviance AUC TSS n_pres 
1 2013 49.84132 0.7978516 0.5000000 182 
2 2021 48.63107 0.8796296 0.6944444 178 
3 2017 41.58266 0.7777778 0.5000000 208 
4 2014 50.67953 0.7706612 0.5000000 192 
5 2022 44.32915 0.8567901 0.5111111 169 
6 2015 44.56914 0.8163265 0.5714286 207 
7 2011 41.00288 0.7159763 0.3846154 201 
8 2020 51.86176 0.9256198 0.7727273 192 
9 2012 47.24728 0.7959184 0.5714286 207 
10 2023 34.93258 0.6250000 0.3750000 206 
11 2018 46.10166 0.7777778 0.5000000 202 
12 2019 46.26296 1.0000000 1.0000000 212 
13 2016 52.58324 0.7500000 0.5000000 212

## Analyzing Response Curves & Variable Importance

Look at Nima's code for pretty plots: <https://github.com/nfarchadi/heatwave_impacts_on_fisheries/blob/main/scripts/4_plots/FS3_FS4_response_relimport_plots.R>

```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models


###----response plots----####

brt_LongfinSmelt <- here("data","LongfinSmelt_brt_SDM_eval_gbm_step.rds") %>% readRDS()


# get the matrix out of the `plot.gbm`
var<-brt_LongfinSmelt$brt$var.names
response_data<-data.frame()

for (i in 1:length(var)){
response_plot_data <- gbm::plot.gbm(brt_LongfinSmelt$brt,
                                 i.var = var[i],
                                 return.grid = TRUE)

response_plot_data<-response_plot_data %>% gather("variable","x",-y)

response_data<-rbind(response_data, response_plot_data)

}

#Get it on the same scale as elith
#response_data$y <- response_data[,2] - mean(response_data[,2])
#response_data$y<-1/(1+exp(-response_data$y))

#response_data$y<-scale(response_data$y, scale = FALSE)

#Add a smoothed line to an active plot
#loess(y ~ x, span = 0.3)

response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size =1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() + 
  theme(panel.spacing = unit(.30, "lines"),
        strip.text = element_text(size=10),
        strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0,0,.05,0, "cm")))
  
ggsave(here("Plots","LongfinSmelt_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)

###----variable contribution----####
rel.inf<-data.frame(Variable = brt_LongfinSmelt$brt$contributions$var,
                        rel.inf=brt_LongfinSmelt$brt$contributions$rel.inf)# %>% 



rel.inf %>% 
  ggplot()+
  geom_bar(aes(x = reorder(Variable, -rel.inf),y=rel.inf), fill = "lightseagreen",
           stat="identity", position=position_dodge())+
  theme_bw() +
  labs(x = "Environmental Variable",
       y = "Relative Importance (%)")+
  theme(axis.text.x = element_text(angle = 45,hjust=1))

ggsave(here("Plots","LongfinSmelt_relative_importance.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","FS3_relative_importance.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```
When looking at partial dependence plots, the y-axis represents the marginal impact of the independent variable to the dependent variable. If at 0, then there is 0 impact...

# NOTES

-   If number of trees below 1k, need to lower learning rate.
-   There may be a difference in the caret output and gbm output (5%). To avoid this can use the default grid to optimize parameters and use predict to have the same results... Or don't worry about it.
