---
title: "Age-0 Longfin Smelt Presence/Absence BRT"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Loading libraries and data

```{r}
library(readr)
library(dismo)
library(caret)
library(here)
library(cowplot)
library(Hmisc)
library(tidyverse)
library(gbm3)
PresAbs<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_Thesis_PresAbs_062524.csv")
```

### Quick Plot of plus counts in data
```{r}
OneFishOneRow<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_OneFishOneRow.csv")
LFS<-OneFishOneRow %>% filter(gensp %in% "Spirinchus thaleichthys")
LFS<-LFS %>% mutate(Type= ifelse(is.na(FishID), "Plus", "Measured"))

LFS %>% ggplot(aes(x=StandardLength, fill=Type)) + geom_histogram() + ggtitle("Length distribution of Longfin smelt 2011-2023") + theme(text=element_text(size=20))# + facet_wrap(LFS$Month, scales = "free")

PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", !AgeClass=="NA") %>% ggplot(aes(x=StandardLength,y=Count, fill=AgeClass)) + geom_bar(stat="identity")+ ggtitle("Length distribution of Longfin smelt 2011-2023") +  theme_bw(base_size = 26)

PresAbs %>% mutate(Type= ifelse(is.na(FishID), "Plus", "Measured")) %>% filter(gensp %in% "Spirinchus thaleichthys") %>% ggplot(aes(x=StandardLength,y=Count, fill=Type)) + geom_bar(stat="identity")+ ggtitle("Length distribution of Longfin smelt 2011-2023") + theme(text=element_text(size=20))

PresAbs$Month<-as.factor(PresAbs$Month)

PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", AgeClass=="Age-0") %>% group_by(Month) %>% dplyr::summarize(TotalCount=sum(Count), TotalTrawlTime=sum(TowDuration), CPUE= TotalCount/TotalTrawlTime) %>% ggplot(aes(x=Month, y=CPUE)) +geom_col()+theme_bw(base_size = 26)+ylab("Catch per minute trawl") +xlab("Month (summed 2011-2023)") + ggtitle("Age-0 longfin smelt CPUE Seasonality")

```

# PresenceAbsence

## Step 1: Wrangling, Visualization, Quality Check

Before modeling, we will ensure the data is up to snuff.

First things first, we need to get the data for the LongfinSmelt of interest we want to model.

### LFS data filtering
```{r}
LongfinSmeltYOYData<-PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", AgeClass %in% "Age-0") %>% dplyr::select(-...1, -...2)
LongfinSmeltYOYData<-as.data.frame(LongfinSmeltYOYData)
unique(LongfinSmeltYOYData$AgeClass)
```

Now lets check the dataset.

```{r}
glimpse(LongfinSmeltYOYData)
summary(LongfinSmeltYOYData)
NumSamples<-unique(LongfinSmeltYOYData$SampleRowID) #3505 sampling events.
```

### 1 row for each sampling event.
```{r}
LongfinSmeltYOYDataPA<-LongfinSmeltYOYData %>% dplyr::group_by(SampleRowID, WaterTemperature, DO, Secchi, Salinity, Year, Month, StationCode) %>% dplyr::summarize(Count=sum(Count)) %>% mutate(PresAbs = if_else(Count > 0, "present", "absent"), Binary=if_else(Count > 0, 1, 0))

LongfinSmeltYOYDataPA$PresAbs<-as.factor(LongfinSmeltYOYDataPA$PresAbs)

LongfinSmeltYOYDataPA$Year<-as.factor(LongfinSmeltYOYDataPA$Year)

LongfinSmeltYOYDataPA$Month<-as.factor(LongfinSmeltYOYDataPA$Month)

LongfinSmeltYOYDataPA$StationCode<-as.factor(LongfinSmeltYOYDataPA$StationCode)

LongfinSmeltYOYDataPA<-as.data.frame(LongfinSmeltYOYDataPA)
```


### Summaries of catch
```{r}
LongfinSmeltYOYDataPA %>% group_by(Year) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) #Boom and bust years, not great for loo validation.

LongfinSmeltYOYDataPA %>% group_by(StationCode) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) %>% print(n=24) #Some are heavily favored.

LongfinSmeltYOYDataPA %>% group_by(Month) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) #Never in January and February

#Overall capture probability
(sum(LongfinSmeltYOYDataPA$Binary)/nrow(LongfinSmeltYOYDataPA))*100
```

6.22% probability of capturing yoy lfs.

### Summaries of samples
```{r}
range(LongfinSmeltYOYDataPA$WaterTemperature)
mean(LongfinSmeltYOYDataPA$WaterTemperature)

range(LongfinSmeltYOYDataPA$DO)
mean(LongfinSmeltYOYDataPA$DO)

range(LongfinSmeltYOYDataPA$Salinity)
mean(LongfinSmeltYOYDataPA$Salinity)

range(LongfinSmeltYOYDataPA$Secchi)
mean(LongfinSmeltYOYDataPA$Secchi)

correlationdf<-LongfinSmeltYOYDataPA %>% select(WaterTemperature, Salinity, DO, Secchi)
cor(correlationdf, method = 'spearman')
```

| Water Quality     | Mean      | Range           |
|-------------------|-----------|-----------------|
| Water temperature | 16.42 C   | 4.6, 27 C       |
| Dissolved oxygen  | 6.97 mg/L | 0.7, 13.74 mg/L |
| Salinity          | 4.69 psu  | 0, 15.8 psu     |
| Secchi            | 30.66 cm  | 7, 99 cm        |

### Visualizations
```{r}
# I want to show presences vs. all samples.
#cdplot(y=LongfinSmeltYOYDataPA$PresAbs, x=LongfinSmeltYOYDataPA$WaterTemperature)

Plot_Water<-LongfinSmeltYOYDataPA %>% ggplot(aes(y=PresAbs, x=WaterTemperature)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw() + ylab("") + xlab("Water Temperature (C)") #+ facet_wrap(~Year)

Plot_DO<-LongfinSmeltYOYDataPA %>% ggplot(aes(y=PresAbs, x=DO)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("")+ xlab("Dissolved Oxygen (mg/L)") #+ facet_wrap(~Year)

Plot_Secchi<-LongfinSmeltYOYDataPA %>% ggplot(aes(y=PresAbs, x=Secchi)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Secchi (cm)") #+ facet_wrap(~Year)

Plot_Salinity<-LongfinSmeltYOYDataPA %>% ggplot(aes(y=PresAbs, x=Salinity)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Salinity") #+ facet_wrap(~Year)

Plots<-plot_grid(Plot_Water, Plot_DO, Plot_Secchi, Plot_Salinity, ncol = 2, labels = "AUTO")

# now add the title
title <- ggdraw() + draw_label("Age-0 Longfin smelt presence/absence by water quality variables for SMFS 2011-2023", fontface = 'bold', x = 0, hjust = 0) +  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7))

plot_grid(title, Plots, ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1))

SummTable <- LongfinSmeltYOYDataPA %>% 
  group_by(PresAbs) %>% #break down yearly summaries by including Year as a grouping variable. 
  dplyr::summarize(
    Mean_WT = mean(WaterTemperature, na.rm=TRUE),
    SD_WT = sd(WaterTemperature, na.rm=TRUE),
    MinWT=min(WaterTemperature, na.rm=TRUE),
    MaxWT=max(WaterTemperature, na.rm=TRUE),
    Mean_DO = mean(DO, na.rm=TRUE),
    SD_DO = sd(DO, na.rm=TRUE),
    MinDO=min(DO, na.rm=TRUE),
    MaxDO=max(DO, na.rm=TRUE),
    Mean_Secchi = mean(Secchi, na.rm=TRUE),
    SD_Secchi = sd(Secchi, na.rm=TRUE),
    MinSecchi=min(Secchi, na.rm=TRUE),
    MaxSecchi=max(Secchi, na.rm=TRUE),
    Mean_Salinity = mean(Salinity, na.rm=TRUE),
    SD_Salinity = sd(Salinity, na.rm=TRUE),
    MinSal=min(Salinity, na.rm=TRUE),
    MaxSal=max(Salinity, na.rm=TRUE)
  ) 

print(SummTable)
write.csv(SummTable, here("data", "LongfinSmeltYOY_SummTable.csv"))

#Seasonality
LongfinSmeltYOYDataPA %>% dplyr::summarize(TotalPresence=sum(Binary, na.rm=TRUE), sd_binary=sd(Binary, na.rm=TRUE), n=n(), PercentageOfPresence=100*(TotalPresence/n)) #Total presence = 1971 of 5471. 36.03% presence of samples.

LongfinSmeltYOYDataPA %>% group_by(Month) %>% dplyr::summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=Month, y=TotalPresence)) +geom_col()+geom_linerange(aes(ymin=TotalPresence-sd_binary, ymax=TotalPresence+sd_binary))+theme_bw(base_size = 26)+ylab("Trawls that catch Age-0 longfin smelt") +xlab("Month (summed 2011-2023)") #+ theme(axis.text.x = element_text(angle = 45, hjust=1))


LongfinSmeltYOYDataPA %>% group_by(StationCode) %>% dplyr::summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=StationCode, y=PercentageOfPresence)) +geom_col()+theme_bw(base_size = 26)+ylab("% of trawls that catch Age-0 longfin smelt") +xlab("Station Code") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Most restricted presence compared to absence is dissolved oxygen. High secchi depth (clear water) also lacks presence above 75 cm.

Do the most influential?

## Step 2: Use gbm() from dismo to run model and create outputs

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

### Setting hyperparameters

```{r}
str(LongfinSmeltYOYDataPA)

 set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
Complexity<-4
LearningRate<-0.005
Bag<-0.6  

```

### gbm.step
```{r}
set.seed(124)
system.time( brt1 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, # Binary response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brt1)

mean(brt1$residuals^2) #mean residual deviance 0.309
#brt1$
names(brt1) #tells us the components of output
CV_deviance_explained <- (((brt1$self.statistics$mean.null- brt1$cv.statistics$deviance.mean)/brt1$self.statistics$mean.null)*100) 

CV_Correlation<-brt1$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)
length(brt1$fitted)#3505

 # reponse curves.
# The difference is that gbm.plot uses the ordinary logit as the outcome for the plots, whereas partial uses a class-centered logit, this choice was made so that the y-axis has the same scale in binomial and multinomial models.
dismo::gbm.plot(brt1, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.
brt1$fit #The fitted values for the 3505 rows.
gbm::plot.gbm(brt1, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brt1)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brt1)
find.int#Salinity and DO, Salinity and Water Temp.
brt1$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brt1$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brt1,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brt1,1, 4) #salinity and DO
gbm.perspec(brt1,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brt1, here("data", "YOYLongfinSmelt_brt_4_005_6_gbm_step.rds"))
```

## Step 3: Validation

### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 1850,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

set.seed(124) #setting the seed for reproducibility. BRT have randomness in them so if you want to make sure you are getting the same result each time make sure to set the seed

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
set.seed(124)
inTrain<-createDataPartition(y=LongfinSmeltYOYDataPA$Binary, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
LongfinSmeltYOYDataPA_Train<-LongfinSmeltYOYDataPA[inTrain,]
LongfinSmeltYOYDataPA_Test<-LongfinSmeltYOYDataPA[-inTrain,]
```

### Running model again

```{r}
set.seed(124)
#LearningRate<-0.001 #had to shift learning rate since only model 80% of the data.
brt2 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA_Train, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, 
                         family = "bernoulli", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees
CV_deviance_explained <- (((brt2$self.statistics$mean.null- brt2$cv.statistics$deviance.mean)/brt2$self.statistics$mean.null)*100) 

brt2$cv.statistics

Preds<-predict(object = brt2, newdata = LongfinSmeltYOYDataPA_Test, type = "response")

#PredGrid<-gbm.predict.grids(brt2, LongfinSmeltYOYDataPA_Test, want.grids=F, n.trees=brt2$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.1,1,0))

LongfinSmeltYOYDataPA_Test$Binary<-as.factor(LongfinSmeltYOYDataPA_Test$Binary)

confusionMatrix(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)

LongfinSmeltYOYDataPA_Test$Binary<-as.character(LongfinSmeltYOYDataPA_Test$Binary)
LongfinSmeltYOYDataPA_Test$Binary<-as.numeric(LongfinSmeltYOYDataPA_Test$Binary)

calc.deviance(LongfinSmeltYOYDataPA_Test$Binary, Preds, calc.mean=T, family="bernoulli") 
library(gbm)

roc(LongfinSmeltYOYDataPA_Test$Binary, Preds) #Area under the curve: 0.8112

ROC<-pROC::roc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
auc<-round(auc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()
PredBinaries<-as.character(PredBinaries)
PredBinaries<-as.numeric(PredBinaries)
cor(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
plot(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/

### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function
set.seed(124)

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYDataPA,
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function
set.seed(124)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 1),],
                        abs = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval1 <- list(brt = brt1, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval1
#now save it

saveRDS(eval, here("data","YOYLongfinSmelt_brt_SDM_eval_gbm_step.rds"))

#leave one water year out function
#To do a loo by water year, I need to add a column for water year to the dataset.
set.seed(124)

LFS_WaterYears<-LongfinSmeltYOYDataPA
#I have to convert year and month to be numeric as they are currently factors. I can't just do as.numeric. I first have to convert them to characters otherwise the factor level will be the number.
LFS_WaterYears$Year<-as.character(LFS_WaterYears$Year)
LFS_WaterYears$Year<-as.numeric(LFS_WaterYears$Year)
LFS_WaterYears$Month<-as.character(LFS_WaterYears$Month)
LFS_WaterYears$Month<-as.numeric(LFS_WaterYears$Month)

LFS_WaterYears$WaterYear<-ifelse(LFS_WaterYears$Month >= 10, LFS_WaterYears$Year + 1, LFS_WaterYears$Year)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$WaterYear
  abs$year <- abs$WaterYear
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LFS_WaterYears[which(LFS_WaterYears$Binary == 1),],
                        abs = LFS_WaterYears[which(LFS_WaterYears$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval2 <- list(brt = brt1, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval2
```


## Step 4: Analyzing Response Curves & Variable Importance

Look at Nima's code for pretty plots: <https://github.com/nfarchadi/heatwave_impacts_on_fisheries/blob/main/scripts/4_plots/FS3_FS4_response_relimport_plots.R>

### Nima plots
```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models

gbm::plot.gbm(brt1, type="response") #Returns a vector of predictions. By default the predictions are on the scale of f(x). For example, for the Bernoulli loss the returned value is on the log odds scale, poisson loss on the log scale, and coxph is on the log hazard scale. If type="response" then gbm converts back to the same scale as the outcome. Currently the only effect this will have is returning probabilities for bernoulli and expected counts for poisson. For the other distributions "response" and "link" return the same.

###----response plots----####

# get the matrix out of the `plot.gbm`
var<-brt1$var.names
response_data<-data.frame()

for (i in 1:length(var)){
response_plot_data <- gbm::plot.gbm(brt1,
                                    i.var = var[i],
                                 return.grid = TRUE)

response_plot_data<-response_plot_data %>% gather("variable","x",-y)

response_data<-rbind(response_data, response_plot_data)

}

#Get it on the same scale as elith
response_data$y <- response_data[,1] - mean(response_data[,1]) #The code above makes the first column y values. This line of code centers the values of the first column around 0.

str(response_data)
response_data<-response_data %>% distinct()
response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() 


ggsave(here("Plots","LongfinSmelt_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)

###----variable contribution----####
rel.inf<-data.frame(Variable = brt_LongfinSmelt$brt$contributions$var,
                        rel.inf=brt_LongfinSmelt$brt$contributions$rel.inf)# %>% 

rel.inf %>% 
  ggplot()+
  geom_bar(aes(x = reorder(Variable, -rel.inf),y=rel.inf), fill = "lightseagreen",
           stat="identity", position=position_dodge())+
  theme_bw() +
  labs(x = "Environmental Variable",
       y = "Relative Importance (%)")+
  theme(axis.text.x = element_text(angle = 45,hjust=1))

ggsave(here("Plots","LongfinSmelt_relative_importance.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","FS3_relative_importance.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```

When looking at partial dependence plots, the y-axis represents the marginal impact of the independent variable to the dependent variable. If at 0, then there is 0 impact...


### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 1100,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

set.seed(124) #setting the seed for reproducibility. BRT have randomness in them so if you want to make sure you are getting the same result each time make sure to set the seed

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

Stochastic Gradient Boosting 3505 samples   4 predictor   2 classes: 'absent', 'present' No pre-processingResampling: Cross-Validated (10 fold, repeated 3 times) Summary of sample sizes: 3155, 3154, 3155, 3154, 3155, 3155, ... Resampling results:  Accuracy  Kappa      0.932194  0.1785341Tuning parameter 'n.trees' was held constant at a value of a value of 0.08Tuning parameter 'n.minobsinnode' was held constant at a value of 10

Stochastic Gradient Boosting 

3505 samples
   4 predictor
   2 classes: 'absent', 'present' 

No pre-processing
Resampling: Cross-Validated (10 fold) 
Summary of sample sizes: 3154, 3155, 3155, 3155, 3155, 3154, ... 
Resampling results:

  ROC       Sens       Spec     
  0.774977  0.9838776  0.1422078

Tuning parameter 'n.trees' was held constant at a value of
 a value of 0.08
Tuning parameter 'n.minobsinnode' was
 held constant at a value of 10
 
## Differing complexity
### TC=1
#### BRT

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

##### Setting hyperparameters

```{r}
str(LongfinSmeltYOYDataPA)

 set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
Complexity<-1
LearningRate<-0.005
Bag<-0.6  

```

##### gbm.step
```{r}
set.seed(124)
system.time( brtc1 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, # Binary response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brtc1)

#mean(brtc1$residuals^2) #mean residual deviance 
names(brtc1) #tells us the components of output
CV_deviance_explained <- (((brtc1$self.statistics$mean.null- brtc1$cv.statistics$deviance.mean)/brtc1$self.statistics$mean.null)*100) 

#CV_Correlation<-brtc1$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)
length(brtc1$fitted)#3505

 # reponse curves.
# The difference is that gbm.plot uses the ordinary logit as the outcome for the plots, whereas partial uses a class-centered logit, this choice was made so that the y-axis has the same scale in binomial and multinomial models.
dismo::gbm.plot(brtc1, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.
brt1$fit #The fitted values for the 3505 rows.
gbm::plot.gbm(brtc1, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brtc1)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brtc1)
find.int#Salinity and DO, Salinity and Water Temp.
#brtc1$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brtc1$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brtc1,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brtc1,1, 4) #salinity and DO
gbm.perspec(brtc1,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brtc1, here("data", "YOYLongfinSmelt_brt_1_005_6_gbm_step.rds"))
```

#### Validation

##### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 3000,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

##### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
set.seed(124)
inTrain<-createDataPartition(y=LongfinSmeltYOYDataPA$Binary, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
LongfinSmeltYOYDataPA_Train<-LongfinSmeltYOYDataPA[inTrain,]
LongfinSmeltYOYDataPA_Test<-LongfinSmeltYOYDataPA[-inTrain,]
```

Running model again

```{r}
set.seed(124)
#LearningRate<-0.001 #had to shift learning rate since only model 80% of the data.
brtc1_2 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA_Train, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, 
                         family = "bernoulli", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees
CV_deviance_explained <- (((brtc1_2$self.statistics$mean.null- brtc1_2$cv.statistics$deviance.mean)/brtc1_2$self.statistics$mean.null)*100) 
brtc1_2$cv.statistics

Preds<-predict(object = brtc1_2, newdata = LongfinSmeltYOYDataPA_Test, type = "response")

#PredGrid<-gbm.predict.grids(brt2, LongfinSmeltYOYDataPA_Test, want.grids=F, n.trees=brt2$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.1,1,0))

LongfinSmeltYOYDataPA_Test$Binary<-as.factor(LongfinSmeltYOYDataPA_Test$Binary)

confusionMatrix(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)

LongfinSmeltYOYDataPA_Test$Binary<-as.character(LongfinSmeltYOYDataPA_Test$Binary)
LongfinSmeltYOYDataPA_Test$Binary<-as.numeric(LongfinSmeltYOYDataPA_Test$Binary)

calc.deviance(LongfinSmeltYOYDataPA_Test$Binary, Preds, calc.mean=T, family="bernoulli") 
library(gbm)

roc(LongfinSmeltYOYDataPA_Test$Binary, Preds) #Area under the curve: 0.8112

ROC<-pROC::roc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
auc<-round(auc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()

PredBinaries<-as.character(PredBinaries)
PredBinaries<-as.numeric(PredBinaries)
cor(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
plot(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/

##### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function
set.seed(124)

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYDataPA,
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function
set.seed(124)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 1),],
                        abs = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval1 <- list(brt = brtc1, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval1
#now save it

#saveRDS(eval, here("data","YOYLongfinSmelt_brt_SDM_eval_gbm_step.rds"))
```

```{r}
#leave one water year out function
#To do a loo by water year, I need to add a column for water year to the dataset.
set.seed(124)

LFS_WaterYears<-LongfinSmeltYOYDataPA
#I have to convert year and month to be numeric as they are currently factors. I can't just do as.numeric. I first have to convert them to characters otherwise the factor level will be the number.
LFS_WaterYears$Year<-as.character(LFS_WaterYears$Year)
LFS_WaterYears$Year<-as.numeric(LFS_WaterYears$Year)
LFS_WaterYears$Month<-as.character(LFS_WaterYears$Month)
LFS_WaterYears$Month<-as.numeric(LFS_WaterYears$Month)

LFS_WaterYears$WaterYear<-ifelse(LFS_WaterYears$Month >= 10, LFS_WaterYears$Year + 1, LFS_WaterYears$Year)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$WaterYear
  abs$year <- abs$WaterYear
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LFS_WaterYears[which(LFS_WaterYears$Binary == 1),],
                        abs = LFS_WaterYears[which(LFS_WaterYears$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval2 <- list(brt = brtc1, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval2
```

### TC = 2
#### BRT

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

##### Setting hyperparameters

```{r}
str(LongfinSmeltYOYDataPA)

 set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
Complexity<-2
LearningRate<-0.005
Bag<-0.6  

```

##### gbm.step
```{r}
set.seed(124)
system.time( brtc2 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, # Binary response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brtc2)

mean(brtc2$residuals^2) #mean residual deviance 0.309
names(brtc2) #tells us the components of output
CV_deviance_explained <- (((brtc2$self.statistics$mean.null- brtc2$cv.statistics$deviance.mean)/brtc2$self.statistics$mean.null)*100) 

CV_Correlation<-brtc2$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)
length(brtc2$fitted)#3505

 # reponse curves.
# The difference is that gbm.plot uses the ordinary logit as the outcome for the plots, whereas partial uses a class-centered logit, this choice was made so that the y-axis has the same scale in binomial and multinomial models.
dismo::gbm.plot(brtc2, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.
brtc2$fit #The fitted values for the 3505 rows.
gbm::plot.gbm(brtc2, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brtc2)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brtc2)
find.int#Salinity and DO, Salinity and Water Temp.
brtc2$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brtc2$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brtc2,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brtc2,1, 4) #salinity and DO
gbm.perspec(brtc2,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brtc2, here("data", "YOYLongfinSmelt_brt_2_005_6_gbm_step.rds"))
```

#### Validation

##### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 2850,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

##### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
set.seed(124)
inTrain<-createDataPartition(y=LongfinSmeltYOYDataPA$Binary, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
LongfinSmeltYOYDataPA_Train<-LongfinSmeltYOYDataPA[inTrain,]
LongfinSmeltYOYDataPA_Test<-LongfinSmeltYOYDataPA[-inTrain,]
```

Running model again

```{r}
#LearningRate<-0.001 #had to shift learning rate since only model 80% of the data.
brt2tc2 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA_Train, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, 
                         family = "bernoulli", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees
CV_deviance_explained <- (((brt2tc2$self.statistics$mean.null- brt2tc2$cv.statistics$deviance.mean)/brt2tc2$self.statistics$mean.null)*100) 
brt2tc2$cv.statistics

Preds<-predict(object = brt2tc2, newdata = LongfinSmeltYOYDataPA_Test, type = "response")

#PredGrid<-gbm.predict.grids(brt2, LongfinSmeltYOYDataPA_Test, want.grids=F, n.trees=brt2$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.1,1,0))

LongfinSmeltYOYDataPA_Test$Binary<-as.factor(LongfinSmeltYOYDataPA_Test$Binary)

confusionMatrix(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)

LongfinSmeltYOYDataPA_Test$Binary<-as.character(LongfinSmeltYOYDataPA_Test$Binary)
LongfinSmeltYOYDataPA_Test$Binary<-as.numeric(LongfinSmeltYOYDataPA_Test$Binary)

calc.deviance(LongfinSmeltYOYDataPA_Test$Binary, Preds, calc.mean=T, family="bernoulli") #6.890836
library(gbm)

roc(LongfinSmeltYOYDataPA_Test$Binary, Preds) #Area under the curve: 0.8112

ROC<-pROC::roc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
auc<-round(auc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()

PredBinaries<-as.character(PredBinaries)
PredBinaries<-as.numeric(PredBinaries)

cor(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
plot(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/

##### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function
set.seed(124)

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYDataPA,
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function
set.seed(124)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 1),],
                        abs = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval1 <- list(brt = brtc2, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval1
#now save it

#saveRDS(eval, here("data","YOYLongfinSmelt_brtc1_SDM_eval_gbm_step.rds"))
```

```{r}
#leave one water year out function
#To do a loo by water year, I need to add a column for water year to the dataset.
set.seed(124)

LFS_WaterYears<-LongfinSmeltYOYDataPA
#I have to convert year and month to be numeric as they are currently factors. I can't just do as.numeric. I first have to convert them to characters otherwise the factor level will be the number.
LFS_WaterYears$Year<-as.character(LFS_WaterYears$Year)
LFS_WaterYears$Year<-as.numeric(LFS_WaterYears$Year)
LFS_WaterYears$Month<-as.character(LFS_WaterYears$Month)
LFS_WaterYears$Month<-as.numeric(LFS_WaterYears$Month)

LFS_WaterYears$WaterYear<-ifelse(LFS_WaterYears$Month >= 10, LFS_WaterYears$Year + 1, LFS_WaterYears$Year)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$WaterYear
  abs$year <- abs$WaterYear
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LFS_WaterYears[which(LFS_WaterYears$Binary == 1),],
                        abs = LFS_WaterYears[which(LFS_WaterYears$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval2 <- list(brt = brtc2, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval2
```
### TC = 3
#### BRT

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

##### Setting hyperparameters

```{r}
str(LongfinSmeltYOYDataPA)

 set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
Complexity<-3
LearningRate<-0.005
Bag<-0.6  

```

##### gbm.step
```{r}
set.seed(124)
system.time( brtc3 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, # Binary response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brtc3)

mean(brtc3$residuals^2) #mean residual deviance 0.309
#brt1$
names(brtc3) #tells us the components of output
CV_deviance_explained <- (((brtc3$self.statistics$mean.null- brtc3$cv.statistics$deviance.mean)/brtc3$self.statistics$mean.null)*100) 

CV_Correlation<-brtc3$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)
length(brtc3$fitted)#3505

 # reponse curves.
# The difference is that gbm.plot uses the ordinary logit as the outcome for the plots, whereas partial uses a class-centered logit, this choice was made so that the y-axis has the same scale in binomial and multinomial models.
dismo::gbm.plot(brtc3, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.
brtc3$fit #The fitted values for the 3505 rows.
gbm::plot.gbm(brtc3, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brtc3)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brtc3)
find.int#Salinity and DO, Salinity and Water Temp.
brtc3$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brtc3$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brtc3,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brtc3,1, 4) #salinity and DO
gbm.perspec(brtc3,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brtc3, here("data", "YOYLongfinSmelt_brt_3_005_6_gbm_step.rds"))
```

#### Validation

##### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 2150,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )


gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

##### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
set.seed(124)
inTrain<-createDataPartition(y=LongfinSmeltYOYDataPA$Binary, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
LongfinSmeltYOYDataPA_Train<-LongfinSmeltYOYDataPA[inTrain,]
LongfinSmeltYOYDataPA_Test<-LongfinSmeltYOYDataPA[-inTrain,]
```

Running model again

```{r}
#LearningRate<-0.001 #had to shift learning rate since only model 80% of the data.
brtc3_2 <- dismo::gbm.step(data=LongfinSmeltYOYDataPA_Train, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 11, 
                         family = "bernoulli", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees
CV_deviance_explained <- (((brtc3_2$self.statistics$mean.null- brtc3_2$cv.statistics$deviance.mean)/brtc3_2$self.statistics$mean.null)*100) 
brtc3_2$cv.statistics

Preds<-predict(object = brtc3_2, newdata = LongfinSmeltYOYDataPA_Test, type = "response")

#PredGrid<-gbm.predict.grids(brt2, LongfinSmeltYOYDataPA_Test, want.grids=F, n.trees=brt2$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.1,1,0))

LongfinSmeltYOYDataPA_Test$Binary<-as.factor(LongfinSmeltYOYDataPA_Test$Binary)

confusionMatrix(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)

LongfinSmeltYOYDataPA_Test$Binary<-as.character(LongfinSmeltYOYDataPA_Test$Binary)

LongfinSmeltYOYDataPA_Test$Binary<-as.numeric(LongfinSmeltYOYDataPA_Test$Binary)

calc.deviance(LongfinSmeltYOYDataPA_Test$Binary, Preds, calc.mean=T, family="bernoulli") 
library(gbm)

roc(LongfinSmeltYOYDataPA_Test$Binary, Preds) #Area under the curve: 0.8112

ROC<-pROC::roc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
auc<-round(auc(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()
PredBinaries<-as.character(PredBinaries)
PredBinaries<-as.numeric(PredBinaries)
cor(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
plot(PredBinaries, LongfinSmeltYOYDataPA_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/

##### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function
set.seed(124)

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYDataPA,
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function
set.seed(124)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 1),],
                        abs = LongfinSmeltYOYDataPA[which(LongfinSmeltYOYDataPA$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval1 <- list(brt = brtc3, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval1
#now save it

#saveRDS(eval, here("data","YOYLongfinSmelt_brt_SDM_eval_gbm_step.rds"))
```

```{r}
#leave one water year out function
#To do a loo by water year, I need to add a column for water year to the dataset.
set.seed(124)

LFS_WaterYears<-LongfinSmeltYOYDataPA
#I have to convert year and month to be numeric as they are currently factors. I can't just do as.numeric. I first have to convert them to characters otherwise the factor level will be the number.
LFS_WaterYears$Year<-as.character(LFS_WaterYears$Year)
LFS_WaterYears$Year<-as.numeric(LFS_WaterYears$Year)
LFS_WaterYears$Month<-as.character(LFS_WaterYears$Month)
LFS_WaterYears$Month<-as.numeric(LFS_WaterYears$Month)

LFS_WaterYears$WaterYear<-ifelse(LFS_WaterYears$Month >= 10, LFS_WaterYears$Year + 1, LFS_WaterYears$Year)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$WaterYear
  abs$year <- abs$WaterYear
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LFS_WaterYears[which(LFS_WaterYears$Binary == 1),],
                        abs = LFS_WaterYears[which(LFS_WaterYears$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=11, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval2 <- list(brt = brtc3, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval2
```

#### Complexities Nima plot
```{r}
###----response plots----####

# get the matrix out of the `plot.gbm`
var<-brt1$var.names
tc4_response_data<-data.frame()

for (i in 1:length(var)){
tc4_response_plot_data <- gbm::plot.gbm(brt1,
                                    i.var = var[i],
                                 return.grid = TRUE)

tc4_response_plot_data<-tc4_response_plot_data %>% gather("variable","x",-y)

tc4_response_data<-rbind(tc4_response_data, tc4_response_plot_data)

}

#Get it on the same scale as elith https://stats.stackexchange.com/questions/122721/r-partial-dependency-plots-from-gbm-package-values-and-y-axis
#response_data$y <- response_data[,2] - mean(response_data[,2])

tc4_response_data$y <- tc4_response_data[,1] - mean(tc4_response_data[,1])

var<-brtc3$var.names
tc3_response_data<-data.frame()

for (i in 1:length(var)){
tc3_response_plot_data <- gbm::plot.gbm(brtc3,
                                    i.var = var[i],
                                 return.grid = TRUE)

tc3_response_plot_data<-tc3_response_plot_data %>% gather("variable","x",-y)

tc3_response_data<-rbind(tc3_response_data, tc3_response_plot_data)

}

#Elith scale
tc3_response_data$y <- tc3_response_data[,1] - mean(tc3_response_data[,1])

var<-brtc2$var.names
tc2_response_data<-data.frame()

for (i in 1:length(var)){
tc2_response_plot_data <- gbm::plot.gbm(brtc2,
                                    i.var = var[i],
                                 return.grid = TRUE)

tc2_response_plot_data<-tc2_response_plot_data %>% gather("variable","x",-y)

tc2_response_data<-rbind(tc2_response_data, tc2_response_plot_data)

}

#Elith scale
tc2_response_data$y <- tc2_response_data[,1] - mean(tc2_response_data[,1])

var<-brtc1$var.names
tc1_response_data<-data.frame()

for (i in 1:length(var)){
tc1_response_plot_data <- gbm::plot.gbm(brtc1,
                                    i.var = var[i],
                                 return.grid = TRUE)

tc1_response_plot_data<-tc1_response_plot_data %>% gather("variable","x",-y)

tc1_response_data<-rbind(tc1_response_data, tc1_response_plot_data)

}

#Elith scale
tc1_response_data$y <- tc1_response_data[,1] - mean(tc1_response_data[,1])

tc4_response_data$Complexity<-"Four"
tc3_response_data$Complexity<-"Three"
tc2_response_data$Complexity<-"Two"
tc1_response_data$Complexity<-"One"

Combined<-rbind(tc1_response_data, tc2_response_data, tc3_response_data, tc4_response_data)

Combined %>%
  ggplot() + 
  geom_line(aes(x=x, y=y, color=Complexity), size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 4) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() + 
  ggtitle("YOY Longfin Smelt Partial Dependence Plots by Complexity")+ theme(text=element_text(size=20))

ggsave(here("Plots","LongfinSmelt_YOY_3brtresponse_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
```

# COUNT DATA
## Loading libraries and data

```{r}
rm(list=ls())
library(readr)
library(dismo)
library(caret)
library(here)
library(cowplot)
library(Hmisc)
library(tidyverse)
library(gbm3)
PresAbs<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_Thesis_PresAbs_062524.csv")

LongfinSmeltYOYData<-PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", AgeClass %in% "Age-0") %>% dplyr::select(-...1, -...2)
LongfinSmeltYOYData$Month<-as.factor(LongfinSmeltYOYData$Month)

```

### Visualizations
```{r}
summary(LongfinSmeltYOYData)

# I want to show presences vs. all samples.

Plot_Water<-LongfinSmeltYOYData %>% mutate(water_temp_bin = as.factor( round(WaterTemperature * 2) / 2))  %>% 
  group_by(water_temp_bin) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=water_temp_bin, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt & water temperature SMFS 2011-2023", y ="Catch per minute trawl", x="Water Temperature (C)", caption = "Water temperatures rounded to 0.5 then grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin.") + theme(axis.text.x = element_text(angle = 45, hjust=1))
#ggplot(aes(y=Count, x=WaterTemperature)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw() + ylab("Count") + xlab("Water Temperature (C)") #+ facet_wrap(~Year)

Plot_DO<-LongfinSmeltYOYData %>% mutate(do_bin = as.factor(round(DO * 2) / 2))  %>% group_by(do_bin) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=do_bin, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt & dissolved oxygen SMFS 2011-2023", y ="Catch per minute trawl", x="Dissolved Oxygen (mg/L)", caption = "Dissolved oxygen rounded to 0.5 then grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")

#LongfinSmeltYOYData %>% ggplot(aes(y=Count, x=DO)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count")+ xlab("Dissolved Oxygen (mg/L)") #+ facet_wrap(~Year)

Plot_Secchi<-LongfinSmeltYOYData %>% group_by(Secchi) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=Secchi, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt & secchi SMFS 2011-2023", y ="Catch per minute trawl", x="Secchi (cm)", caption = "Secchi grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")
#LongfinSmeltYOYData %>% ggplot(aes(y=Count, x=Secchi)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count") + xlab("Secchi (cm)") #+ facet_wrap(~Year)

Plot_Salinity<-LongfinSmeltYOYData %>% mutate(salinity_bin = as.factor(round(Salinity * 2) / 2))  %>% group_by(salinity_bin) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=salinity_bin, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt & salinity SMFS 2011-2023", y ="Catch per minute trawl", x="Salinity", caption = "Salinity  rounded to 0.5 then grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")
#LongfinSmeltYOYData %>% ggplot(aes(y=Count, x=Salinity))+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count") + xlab("Salinity") #+ facet_wrap(~Year)

Plots<-plot_grid(Plot_Water, Plot_DO, Plot_Secchi, Plot_Salinity, ncol = 2, labels = "AUTO")

# now add the title
title <- ggdraw() + draw_label("Age-0 Longfin smelt count by water quality variables for SMFS 2011-2023", fontface = 'bold', x = 0, hjust = 0) +  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7))

plot_grid(title, Plots, ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1))


#Seasonality
LongfinSmeltYOYData %>% dplyr::summarize(TotalCount=sum(Count), sd_Count=sd(Count), n=n()) #Total presence = 1971 of 5471. 

LongfinSmeltYOYData %>%  group_by(Month) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=Month, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt by month SMFS 2011-2023", y ="Catch per minute trawl", x="Month", caption = "Month grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")
#group_by(Month) %>% dplyr::summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=Month, y=TotalPresence)) +geom_col()+geom_linerange(aes(ymin=TotalPresence-sd_binary, ymax=TotalPresence+sd_binary))+theme_bw(base_size = 26)+ylab("Trawls that catch yoy longfin smelt") +xlab("Month (summed 2011-2023)") #+ theme(axis.text.x = element_text(angle = 45, hjust=1))
LongfinSmeltYOYData %>% group_by(StationCode) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=StationCode, y=CPUE))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "YOY Longfin smelt by station SMFS 2011-2023", y ="Catch per minute trawl", x="Station Code", caption = "Station code grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
  
  #group_by(StationCode) %>% dplyr::summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=StationCode, y=PercentageOfPresence)) +geom_col()+theme_bw(base_size = 26)+ylab("% of trawls that catch yoy longfin smelt") +xlab("Station Code") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```

## LFS data filtering for all counts
```{r}
LongfinSmeltYOYData_All<-PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", AgeClass %in% "Age-0") %>% dplyr::select(-...1, -...2) %>% group_by(SampleRowID, WaterTemperature, Salinity, Secchi, DO) %>% summarize(Count=sum(Count))
LongfinSmeltYOYData_All<-as.data.frame(LongfinSmeltYOYData_All)
```

## Step 2: Use gbm() from dismo to run model and create outputs

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

### Setting hyperparameters

```{r}
set.seed(124) # for reproducibility 
Complexity<-4
LearningRate<-0.005
Bag<-0.6  

```

### gbm.step
```{r}
system.time( brt_allcount <- dismo::gbm.step(data=LongfinSmeltYOYData_All, 
                         gbm.x= c(2,3,4,5), # environmental variables.
                         gbm.y= 6, # Binary response variable
                         family = "poisson", #Or gaussian
                         tree.complexity = Complexity, 
                         learning.rate = LearningRate,  
                         bag.fraction = Bag 
                         ))

#summary(brt_allcount) #relative influence - ignore.

mean(brt_allcount$residuals^2) #mean residual deviance 0..7016824

names(brt_allcount) #tells us the components of output
CV_deviance_explained <- (((brt_allcount$self.statistics$mean.null- brt_allcount$cv.statistics$deviance.mean)/brt_allcount$self.statistics$mean.null)*100) #33.35715

CV_Correlation<-brt_allcount$cv.statistics$correlation.mean #0.3608818

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)


 # reponse curves
dismo::gbm.plot(brt_allcount, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.

gbm::plot.gbm(brt_allcount)#, return.grid=TRUE)

dismo::gbm.plot.fits(brt_allcount)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brt_allcount)
find.int#Salinity and DO, Salinity and Water Temp.
brt_allcount$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brt_allcount$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brt_allcount,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brt_allcount,1, 4) #salinity and DO
gbm.perspec(brt_allcount,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brt_allcount, here("data", "LongfinSmelt_allcount_brt_gbm_step.rds"))
```

## Step 3: Validation And Plots
### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 3950,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling


gbmFit <- train(Count ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYData_All, 
                method = "gbm", 
                trControl = trainControl(method = "cv", number = 10), 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="poisson",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                #metric = "Kappa" #Or ROC 
                )

gbmFit


# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```
### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

NOT BUILT FOR COUNTS 
```{r}
#k fold cross validation function

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="poisson", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYData_All,
                        gbm.x = c(2,3,4,5),
                        gbm.y=6, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

BROKEN: NOT BUILT FOR COUNTS 
```{r}
#leave one year out function not built for count data, unsure how to edit for that!

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="poisson", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYData_All[which(LongfinSmeltYOYData_All$Binary == 1),],
                        abs = LongfinSmeltYOYData_All[which(LongfinSmeltYOYData_All$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=6, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval <- list(brt = brt_allcount, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval
#now save it

saveRDS(eval, here("data","YOYLongfinSmelt_brt_SDM_eval_gbm_step.rds"))
```


## Step 4: Analyzing Response Curves

Look at Nima's code for pretty plots: <https://github.com/nfarchadi/heatwave_impacts_on_fisheries/blob/main/scripts/4_plots/FS3_FS4_response_relimport_plots.R>

### Nima plots
```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models

gbm::plot.gbm(brt_allcount, type="response") #Returns a vector of predictions. By default the predictions are on the scale of f(x). For example, for the Bernoulli loss the returned value is on the log odds scale, poisson loss on the log scale, and coxph is on the log hazard scale. If type="response" then gbm converts back to the same scale as the outcome. Currently the only effect this will have is returning probabilities for bernoulli and expected counts for poisson. For the other distributions "response" and "link" return the same.

###----response plots----####

# get the matrix out of the `plot.gbm`
var<-brt_allcount$var.names
count_response_data<-data.frame()

for (i in 1:length(var)){
count_response_plot_data <- gbm::plot.gbm(brt_allcount,
                                    i.var = var[i],
                                 return.grid = TRUE)

count_response_plot_data<-count_response_plot_data %>% gather("variable","x",-y)

count_response_data<-rbind(count_response_data, count_response_plot_data)

}

#Get it on the same scale as elith https://stats.stackexchange.com/questions/122721/r-partial-dependency-plots-from-gbm-package-values-and-y-axis

count_response_data$y <- count_response_data[,1] - mean(count_response_data[,1])


str(count_response_data)
count_response_data<-count_response_data %>% distinct()
count_response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() 

ggsave(here("Plots","LongfinSmelt_count_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```


## LFS data filtering for above 0 counts
```{r}
LongfinSmeltYOYData_Positive<-PresAbs %>% filter(gensp %in% "Spirinchus thaleichthys", AgeClass %in% "Age-0", Binary == 1) %>% dplyr::select(-...1, -...2) %>% group_by(SampleRowID, WaterTemperature, Salinity, Secchi, DO) %>% summarize(Count=sum(Count))
LongfinSmeltYOYData_Positive<-as.data.frame(LongfinSmeltYOYData_Positive)
unique(LongfinSmeltYOYData_Positive$AgeClass)
summary(LongfinSmeltYOYData_Positive)
```


## Step 2: Use gbm() from dismo to run model and create outputs

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

### Setting hyperparameters

```{r}
str(LongfinSmeltYOYData_Positive)

 set.seed(124) # for reproducibility 
Complexity<-4
LearningRate<-0.001
Bag<-0.6  

```

### gbm.step
```{r}
system.time( brt_pcount <- dismo::gbm.step(data=LongfinSmeltYOYData_Positive, 
                         gbm.x= c(2,3,4,5), # environmental variables.
                         gbm.y= 6, # Binary response variable
                         family = "poisson", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brt_pcount)

mean(brt_pcount$residuals^2) #mean residual deviance 0.309
#brt1$
names(brt_pcount) #tells us the components of output
CV_deviance_explained <- (((brt_pcount$self.statistics$mean.null- brt_pcount$cv.statistics$deviance.mean)/brt_pcount$self.statistics$mean.null)*100) 

CV_Correlation<-brt_pcount$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)


 # reponse curves
dismo::gbm.plot(brt_pcount, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.

gbm::plot.gbm(brt_pcount, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brt_pcount)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brt_pcount)
find.int#Salinity and DO, Salinity and Water Temp.
brt_pcount$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brt_pcount$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brt_pcount,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brt_pcount,1, 4) #salinity and DO
gbm.perspec(brt_pcount,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brt_pcount, here("data", "LongfinSmelt_positivecount_brt_gbm_step.rds"))
```

## Step 3: Validation

### Looking at caret outputs for optimal model (RMSE, Rsquared, MAE)

```{r}
set.seed(124)
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 2950,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling


gbmFit <- train(Count ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYData_Positive, 
                method = "gbm", 
                trControl = trainControl(method = "cv", number = 10), 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="poisson",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                #metric = "Kappa" #Or ROC 
                )

gbmFit


# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

## Step 4: Analyzing Response Curves & Variable Importance

Look at Nima's code for pretty plots: <https://github.com/nfarchadi/heatwave_impacts_on_fisheries/blob/main/scripts/4_plots/FS3_FS4_response_relimport_plots.R>

### Nima plots
```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models

gbm::plot.gbm(brt_pcount, type="response") #Returns a vector of predictions. By default the predictions are on the scale of f(x). For example, for the Bernoulli loss the returned value is on the log odds scale, poisson loss on the log scale, and coxph is on the log hazard scale. If type="response" then gbm converts back to the same scale as the outcome. Currently the only effect this will have is returning probabilities for bernoulli and expected counts for poisson. For the other distributions "response" and "link" return the same.

###----response plots----####

# get the matrix out of the `plot.gbm`
var<-brt_pcount$var.names
pcount_response_data<-data.frame()

for (i in 1:length(var)){
pcount_response_plot_data <- gbm::plot.gbm(brt_pcount,
                                    i.var = var[i],
                                 return.grid = TRUE)

pcount_response_plot_data<-pcount_response_plot_data %>% gather("variable","x",-y)

pcount_response_data<-rbind(pcount_response_data, pcount_response_plot_data)

}

#Get it on the same scale as elith https://stats.stackexchange.com/questions/122721/r-partial-dependency-plots-from-gbm-package-values-and-y-axis
#response_data$y <- response_data[,2] - mean(response_data[,2])

pcount_response_data$y <- pcount_response_data[,1] - mean(pcount_response_data[,1])


str(count_response_data)
pcount_response_data<-pcount_response_data %>% distinct()
pcount_response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() + 
  theme(panel.spacing = unit(.30, "lines"),
        strip.text = element_text(size=10),
        strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0,0,.05,0, "cm")))#+
  #loess(y ~ x, span = 0.3)#Add a smoothed line to an active plot

ggsave(here("Plots","LongfinSmelt_count_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```

```{r}
#Both counts and binomial on same plot...

count_response_data$Model<-"All Count"
pcount_response_data$Model<-"Positive Count"
response_data$Model<-"Presence/Absence"

Combined<-rbind(count_response_data, response_data, pcount_response_data)

Combined %>%
  ggplot() + 
  geom_line(aes(x=x, y=y, color=Model), size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 4) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() + 
  ggtitle("YOY Longfin Smelt Partial Dependence Plots by Response")

ggsave(here("Plots","LongfinSmelt_YOY_3brtresponse_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
```

When looking at partial dependence plots, the y-axis represents the marginal impact of the independent variable to the dependent variable. If at 0, then there is 0 impact...


### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 1100,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

set.seed(124) #setting the seed for reproducibility. BRT have randomness in them so if you want to make sure you are getting the same result each time make sure to set the seed

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = LongfinSmeltYOYDataPA, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

Stochastic Gradient Boosting 3505 samples   4 predictor   2 classes: 'absent', 'present' No pre-processingResampling: Cross-Validated (10 fold, repeated 3 times) Summary of sample sizes: 3155, 3154, 3155, 3154, 3155, 3155, ... Resampling results:  Accuracy  Kappa      0.932194  0.1785341Tuning parameter 'n.trees' was held constant at a value of a value of 0.08Tuning parameter 'n.minobsinnode' was held constant at a value of 10

Stochastic Gradient Boosting 

3505 samples
   4 predictor
   2 classes: 'absent', 'present' 

No pre-processing
Resampling: Cross-Validated (10 fold) 
Summary of sample sizes: 3154, 3155, 3155, 3155, 3155, 3154, ... 
Resampling results:

  ROC       Sens       Spec     
  0.774977  0.9838776  0.1422078

Tuning parameter 'n.trees' was held constant at a value of
 a value of 0.08
Tuning parameter 'n.minobsinnode' was
 held constant at a value of 10
 

 
# NOTES

-   If number of trees below 1k, need to lower learning rate.
-   There may be a difference in the caret output and gbm output (5%). To avoid this can use the default grid to optimize parameters and use predict to have the same results... Or don't worry about it.

