---
title: "Modeling Protocol"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Loading libraries and data

```{r}
library(readr)
library(dismo)
library(caret)
library(here)
library(cowplot)
library(Hmisc)
library(tidyverse)
library(gbm3)
PresAbs<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_Thesis_PresAbs_062524.csv")
```


# Data visualization, Quality Check

## PricklySculpin data filtering and queries
```{r}
PricklySculpinData<-PresAbs %>% filter(gensp %in% "Cottus asper") %>% dplyr::select(-...1, -...2)
PricklySculpinData<-as.data.frame(PricklySculpinData)
unique(PricklySculpinData$AgeClass)
```

Ages
```{r}
PricklySculpinData %>% group_by(AgeClass)%>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100)# %>% print(n=26)
```
Now lets check the dataset.

```{r}
glimpse(PricklySculpinData)
summary(PricklySculpinData)
NumSamples<-unique(PricklySculpinData$SampleRowID) #3505 sampling events.
unique(PricklySculpinData$TowDuration)#No NAs for wt, sal, do, sec
```

Let's make a couple of plots for making predictions. First I need to make the data be just 1 row for each sampling event.

# Separate Ages: 1 row for each sampling event for each age.
```{r}
PricklySculpinDataPA<-PricklySculpinData %>% dplyr::group_by(SampleRowID, WaterTemperature, DO, Secchi, Salinity, Year, Month, StationCode, AgeClass) %>% dplyr::summarize(Count=sum(Count)) %>% mutate(PresAbs = if_else(Count > 0, "present", "absent"), Binary=if_else(Count > 0, 1, 0)) #For modeling ages separately

PricklySculpinDataPA$PresAbs<-as.factor(PricklySculpinDataPA$PresAbs)

PricklySculpinDataPA$Year<-as.factor(PricklySculpinDataPA$Year)

PricklySculpinDataPA$Month<-as.factor(PricklySculpinDataPA$Month)

PricklySculpinDataPA$StationCode<-as.factor(PricklySculpinDataPA$StationCode)

PricklySculpinDataPA<-as.data.frame(PricklySculpinDataPA)
```
## PresenceAbsence

### Summaries 

of catch
```{r}
PricklySculpinDataPA %>% group_by(AgeClass) %>% dplyr::summarize(N=length(unique(SampleRowID)), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) #Boom and bust years, not great for loo validation.

print(PricklySculpinDataPA %>% group_by(AgeClass,Year) %>% dplyr::summarize(N=length(unique(SampleRowID)), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100), n=26) #Boom and bust years, not great for loo validation.

PricklySculpinDataPA %>% group_by(StationCode) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) %>% print(n=24) #Some are heavily favored.

PricklySculpinDataPA %>% group_by(Month) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) #Never in January and February

#Overall capture probability
OverallProb<-(sum(PricklySculpinDataPA$Binary)/nrow(PricklySculpinDataPA))*100
```

`OverallProb`% probability of capturing PricklySculpin.

Summaries of samples
These ranges and means are unaffected by having doubled samples (one for each age class). 
```{r} 
RangeWT<-range(PricklySculpinDataPA$WaterTemperature)
MeanWT<-mean(PricklySculpinDataPA$WaterTemperature)

RangeDO<-range(PricklySculpinDataPA$DO)
MeanDO<-mean(PricklySculpinDataPA$DO)

RangeSal<-range(PricklySculpinDataPA$Salinity)
MeanSal<-mean(PricklySculpinDataPA$Salinity)

RangeSec<-range(PricklySculpinDataPA$Secchi)
MeanSec<-mean(PricklySculpinDataPA$Secchi)

```


| Water Quality     | Mean         | Range           |
|-------------------|--------------|-----------------|
| Water temperature |`MeanWT` C    | `RangeWT` C     |
| Dissolved oxygen  |`MeanDO` mg/L | `RangeDO` mg/L  |
| Salinity          |`MeanSal` psu | `RangeSal` psu  |
| Secchi            |`MeanSec` cm  | `RangeSec` cm   |

### Visualizations
```{r}
# I want to show presences vs. all samples. Keep in mind that the ages are separate but included.

Plot_Water<-PricklySculpinDataPA %>% ggplot(aes(y=PresAbs, x=WaterTemperature)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw() + ylab("") + xlab("Water Temperature (C)") #+ facet_wrap(~Year)

Plot_DO<-PricklySculpinDataPA %>% ggplot(aes(y=PresAbs, x=DO)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("")+ xlab("Dissolved Oxygen (mg/L)") #+ facet_wrap(~Year)

Plot_Secchi<-PricklySculpinDataPA %>% ggplot(aes(y=PresAbs, x=Secchi)) + 
  geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Secchi (cm)") #+ facet_wrap(~Year)

Plot_Salinity<-PricklySculpinDataPA %>% ggplot(aes(y=PresAbs, x=Salinity)) + geom_violin(fill = "lightseagreen")+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("") + xlab("Salinity") #+ facet_wrap(~Year)

Plots<-plot_grid(Plot_Water, Plot_DO, Plot_Secchi, Plot_Salinity, ncol = 2, labels = "AUTO")

# now add the title
title <- ggdraw() + draw_label("PricklySculpin presence/absence by water quality variables for SMFS 2011-2023", fontface = 'bold', x = 0, hjust = 0) +  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7))

plot_grid(title, Plots, ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1))

SummTable <- PricklySculpinDataPA %>% #Keep in mind that the ages are separate but included. A sampling event with both adult and age-0 fish count as present for 2 samples.
  group_by(PresAbs) %>% #break down yearly summaries by including Year as a grouping variable. 
  dplyr::summarize(
    Mean_WT = mean(WaterTemperature, na.rm=TRUE),
    SD_WT = sd(WaterTemperature, na.rm=TRUE),
    MinWT=min(WaterTemperature, na.rm=TRUE),
    MaxWT=max(WaterTemperature, na.rm=TRUE),
    Mean_DO = mean(DO, na.rm=TRUE),
    SD_DO = sd(DO, na.rm=TRUE),
    MinDO=min(DO, na.rm=TRUE),
    MaxDO=max(DO, na.rm=TRUE),
    Mean_Secchi = mean(Secchi, na.rm=TRUE),
    SD_Secchi = sd(Secchi, na.rm=TRUE),
    MinSecchi=min(Secchi, na.rm=TRUE),
    MaxSecchi=max(Secchi, na.rm=TRUE),
    Mean_Salinity = mean(Salinity, na.rm=TRUE),
    SD_Salinity = sd(Salinity, na.rm=TRUE),
    MinSal=min(Salinity, na.rm=TRUE),
    MaxSal=max(Salinity, na.rm=TRUE)
  )

print(SummTable)
write.csv(SummTable, here("data", "PricklySculpin_SummTable.csv"))

#Seasonality
PricklySculpinDataPA %>% group_by(AgeClass) %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=length(unique(SampleRowID)), PercentageOfPresence=100*(TotalPresence/n))

PricklySculpinDataPA$Month<-as.factor(PricklySculpinDataPA$Month)

PricklySculpinDataPA %>% group_by(Month, AgeClass) %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=length(unique(SampleRowID)), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=Month, y=TotalPresence, fill=AgeClass)) +geom_col()+geom_linerange(aes(ymin=TotalPresence-sd_binary, ymax=TotalPresence+sd_binary))+theme_bw(base_size = 26)+ylab("Trawls that catch PricklySculpin") +xlab("Month (summed 2011-2023)") #+ theme(axis.text.x = element_text(angle = 45, hjust=1))
PricklySculpinDataPA %>% group_by(StationCode, AgeClass) %>% summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=length(unique(SampleRowID)), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=StationCode, y=PercentageOfPresence, fill=AgeClass)) +geom_col()+theme_bw(base_size = 26)+ylab("% of trawls that catch PricklySculpin") +xlab("Station Code") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```

# Age 0
## Use gbm() from dismo to run model and create outputs


### Setting data and hyperparameters
```{r}
PricklySculpinAge0<-PricklySculpinDataPA %>% filter(AgeClass %in% "Age-0")
str(PricklySculpinAge0)
unique(PricklySculpinAge0$Binary)
#gbm.step wants bernoulli response {0, 1} 
Complexity<-4
LearningRate<-0.005
Bag<-0.6  
```

### gbm.step
Thanks Nima and Elith!

```{r}
set.seed(124) # for reproducibility 
#gbm.step wants bernoulli response {0, 1} 
SCPbrt_Age0 <- dismo::gbm.step(data=PricklySculpinAge0, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 12, # Binary response variable
                         family = "bernoulli", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         )

# make sure to save the model!
summary(SCPbrt_Age0)

names(SCPbrt_Age0) #tells us the components of output
mean(SCPbrt_Age0$residuals^2) #mean residual deviance 
# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)

 # reponse curves
dismo::gbm.plot(SCPbrt_Age0, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value")

gbm::plot.gbm(SCPbrt_Age0, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(SCPbrt_Age0)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.


CV_deviance_explained <- (((SCPbrt_Age0$self.statistics$mean.null- SCPbrt_Age0$cv.statistics$deviance.mean)/SCPbrt_Age0$self.statistics$mean.null)*100)
SCPbrt_Age0$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.
find.int <- gbm.interactions(SCPbrt_Age0)
find.int
SCPbrt_Age0$self.statistics$correlation[[1]]
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.
#Use these for plotting interactions
# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(SCPbrt_Age0,1, 3) #3D PLOT of Water Temp and Secchi
gbm.perspec(SCPbrt_Age0,1, 2) #Water temp and salinity
#Can also rotate by switching x and y.


# save the model
saveRDS(SCPbrt_Age0, here("data", "PricklySculpinAge0_brt_4_005_6_gbm_step.rds"))

```

Firstly, the things you can see: The R console will show some results (see the Word Working Guide of the Elith tutorial for an example). It reports a brief model summary & all the values are also retained in the model object, so they will be permanently kept (as long as you save the R workspace before quitting).

There will also be a graph.. This model was built with the default 10-fold cross-validation (CV). The solid black curve is the mean, and the dotted curves +- 1 standard error, for the changes in predictive deviance (ie as measured on the excluded folds of the CV). The red line shows the minimum of the mean, and the green line the number of trees at which that occurs. The final model that is returned in the model object is built on the full data set, using the number of trees identified as optimal.

## Step 3: Validation

### 10-fold and loo

Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#10 fold cross validation function
set.seed(124)

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="bernoulli", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = PricklySculpinAge0,
                        gbm.x = c(2,3,4,5),
                        gbm.y=12, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function
set.seed(124)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = PricklySculpinAge0[which(PricklySculpinAge0$Binary == 1),],
                        abs = PricklySculpinAge0[which(PricklySculpinAge0$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=12, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval <- list(brt = SCPbrt_Age0, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval
#now save it

saveRDS(eval, here("data","PricklySculpinAge0_calyear_brt_SDM_eval_gbm_step.rds"))

#leave one water year out function
#To do a loo by water year, I need to add a column for water year to the dataset.
set.seed(124)

SCP_WaterYears<-PricklySculpinAge0
#I have to convert year and month to be numeric as they are currently factors. I can't just do as.numeric. I first have to convert them to characters otherwise the factor level will be the number.
SCP_WaterYears$Year<-as.character(SCP_WaterYears$Year)
SCP_WaterYears$Year<-as.numeric(SCP_WaterYears$Year)
SCP_WaterYears$Month<-as.character(SCP_WaterYears$Month)
SCP_WaterYears$Month<-as.numeric(SCP_WaterYears$Month)

SCP_WaterYears$WaterYear<-ifelse(SCP_WaterYears$Month >= 10, SCP_WaterYears$Year + 1, SCP_WaterYears$Year)

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$WaterYear
  abs$year <- abs$WaterYear
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="bernoulli", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = SCP_WaterYears[which(SCP_WaterYears$Binary == 1),],
                        abs = SCP_WaterYears[which(SCP_WaterYears$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=12, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval2 <- list(brt = SCPbrt_Age0, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval2
saveRDS(eval2, here("data","PricklySculpinAge0_watyear_brt_SDM_eval_gbm_step.rds"))
```

# Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 2250,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

set.seed(124) #setting the seed for reproducibility. BRT have randomness in them so if you want to make sure you are getting the same result each time make sure to set the seed

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = PricklySculpinAge0, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = PricklySculpinAge0, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

### Nima plots
```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models

gbm::plot.gbm(SCPbrt_Age0, type="response") #Returns a vector of predictions. By default the predictions are on the scale of f(x). For example, for the Bernoulli loss the returned value is on the log odds scale, poisson loss on the log scale, and coxph is on the log hazard scale. If type="response" then gbm converts back to the same scale as the outcome. Currently the only effect this will have is returning probabilities for bernoulli and expected counts for poisson. For the other distributions "response" and "link" return the same.

###----response plots----####

# get the matrix out of the `plot.gbm`
var<-SCPbrt_Age0$var.names
response_data<-data.frame()

for (i in 1:length(var)){
response_plot_data <- gbm::plot.gbm(SCPbrt_Age0,
                                    i.var = var[i],
                                 return.grid = TRUE)

response_plot_data<-response_plot_data %>% gather("variable","x",-y)

response_data<-rbind(response_data, response_plot_data)

}

#Get it on the same scale as elith
response_data$y <- response_data[,1] - mean(response_data[,1]) #The code above makes the first column y values. This line of code centers the values of the first column around 0.

str(response_data)
response_data<-response_data %>% distinct()
response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size=1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() 


ggsave(here("Plots","PricklySculpin_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```

### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
inTrain<-createDataPartition(y=PricklySculpinAge0$Binary, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
PricklySculpinAge0_Train<-PricklySculpinAge0[inTrain,]
PricklySculpinAge0_Test<-PricklySculpinAge0[-inTrain,]
```

### Running model again

```{r}
set.seed(124)
#LearningRate<-0.001 #had to shift learning rate since only model 80% of the data.
brtest <- dismo::gbm.step(data=PricklySculpinAge0_Train, 
                         gbm.x= c(2,3,4,5), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 12, 
                         family = "bernoulli", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees

CV_deviance_explained <- (((brtest$self.statistics$mean.null- brtest$cv.statistics$deviance.mean)/brtest$self.statistics$mean.null)*100)

brtest$cv.statistics

Preds<-predict(object = brtest, newdata = PricklySculpinAge0_Test, type = "response")

#PredGrid<-gbm.predict.grids(brtest, PricklySculpinAge0_Test, want.grids=F, n.trees=brtest$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.1,1,0))
PredBinaries<-as.character(PredBinaries)
PredBinaries<-as.numeric(PredBinaries)
#PricklySculpinAge0_Test$Binary<-as.factor(PricklySculpinAge0_Test$Binary)

confusionMatrix(PredBinaries, PricklySculpinAge0_Test$Binary)

PricklySculpinAge0_Test$Binary<-as.character(PricklySculpinAge0_Test$Binary)
PricklySculpinAge0_Test$Binary<-as.numeric(PricklySculpinAge0_Test$Binary)

calc.deviance(PricklySculpinAge0_Test$Binary, Preds, calc.mean=T, family="bernoulli") 

roc(PricklySculpinAge0_Test$Binary, Preds) #Area under the curve: 0.8104

ROC<-pROC::roc(PredBinaries, PricklySculpinAge0_Test$Binary)
auc<-round(auc(PredBinaries, PricklySculpinAge0_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()


cor(PredBinaries, PricklySculpinAge0_Test$Binary)
plot(PredBinaries, PricklySculpinAge0_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/


# COUNT DATA
## Loading libraries and data

```{r}
library(readr)
library(dismo)
library(caret)
library(here)
library(cowplot)
library(Hmisc)
library(tidyverse)
library(gbm3)
PresAbs<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_Thesis_PresAbs_062524.csv")
```

### Visualizations
```{r}
SCPData<-PresAbs %>% filter(gensp %in% "Cottus asper") %>% dplyr::select(-...1, -...2)
SCPData$Month<-as.factor(SCPData$Month)

summary(SCPData)

# I want to show presences vs. all samples.

Plot_Water<-SCPData %>% mutate(water_temp_bin = as.factor( round(WaterTemperature * 2) / 2))  %>% 
  group_by(water_temp_bin, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=water_temp_bin, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "Prickly sculpin & water temperature SMFS 2011-2023", y ="Catch per minute trawl", x="Water Temperature (C)", caption = "Water temperatures rounded to 0.5 then grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin.") + theme(axis.text.x = element_text(angle = 45, hjust=1))+ guides(fill="none")
#ggplot(aes(y=Count, x=WaterTemperature)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw() + ylab("Count") + xlab("Water Temperature (C)") #+ facet_wrap(~Year)

Plot_DO<-SCPData %>% mutate(do_bin = as.factor(round(DO * 2) / 2))  %>% group_by(do_bin, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=do_bin, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "Prickly sculpin & dissolved oxygen SMFS 2011-2023", y ="Catch per minute trawl", x="Dissolved Oxygen (mg/L)", caption = "Dissolved oxygen rounded to 0.5  then grouped to get sum of counts and\n sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")

#SCPData %>% ggplot(aes(y=Count, x=DO)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count")+ xlab("Dissolved Oxygen (mg/L)") #+ facet_wrap(~Year)

Plot_Secchi<-SCPData %>% group_by(Secchi, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=Secchi, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "Prickly sculpin & secchi SMFS 2011-2023", y ="Catch per minute trawl", x="Secchi (cm)", caption = "Secchi grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")+ guides(fill="none")
#SCPData %>% ggplot(aes(y=Count, x=Secchi)) + stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count") + xlab("Secchi (cm)") #+ facet_wrap(~Year)

Plot_Salinity<-SCPData %>% mutate(salinity_bin = as.factor(round(Salinity * 2) / 2))  %>% group_by(salinity_bin, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=salinity_bin, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw()+ labs(title = "Prickly sculpin & salinity SMFS 2011-2023", y ="Catch per minute trawl", x="Salinity", caption = "Salinity  rounded to 0.5 then grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin") + guides(fill="none")
#SCPData %>% ggplot(aes(y=Count, x=Salinity))+ stat_summary(fun.data="mean_sdl", fun.args=list(mult=1), geom= "pointrange", color="black") + theme_bw()+ ylab("Count") + xlab("Salinity") #+ facet_wrap(~Year)

Plots<-plot_grid(Plot_Water, Plot_DO, Plot_Secchi, Plot_Salinity, ncol = 2, labels = "AUTO")

# now add the title
title <- ggdraw() + draw_label("Prickly sculpin count by water quality variables for SMFS 2011-2023", fontface = 'bold', x = 0, hjust = 0) +  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7))

plot_grid(title, Plots, ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1))


#Seasonality
SCPData %>% dplyr::summarize(TotalCount=sum(Count), sd_Count=sd(Count), Effort=sum(TowDuration), CPUE=(TotalCount/Effort)) #Total presence = 3455 of 3505 36.03% presence of samples.
SCPData %>%  group_by(Month, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=Month, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw(base_size = 26)+ labs(title = "Prickly sculpin by month SMFS 2011-2023", y ="Catch per minute trawl", x="Month")#, caption = "Month, age class grouped to get sum of counts and sum trawl duration for that month, age class combination. CPUE = counts by month, age class divided by total tow minutes for that month, age class")

SCPData %>% group_by(StationCode, AgeClass) %>% summarize(TotalCount=sum(Count), TotalTow=sum(TowDuration), CPUE = TotalCount/TotalTow) %>% 
  ggplot(aes(x=StationCode, y=CPUE, fill=AgeClass))+
  geom_bar(stat="identity")+ 
  theme_bw(base_size = 22)+ labs(title = "Prickly sculpin by station SMFS 2011-2023", y ="Catch per minute trawl", x="Station Code")+ theme(axis.text.x = element_text(angle = 45, hjust=1))#, caption = "Station code grouped to get sum of counts and sum trawl duration to get CPUE of counts by bin divided by total tow minutes for that bin")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
  
  #group_by(StationCode) %>% dplyr::summarize(TotalPresence=sum(Binary), sd_binary=sd(Binary), n=n(), PercentageOfPresence=100*(TotalPresence/n)) %>% ggplot(aes(x=StationCode, y=PercentageOfPresence)) +geom_col()+theme_bw(base_size = 26)+ylab("% of trawls that catch yoy longfin smelt") +xlab("Station Code") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```

## Data filtering for ALL counts
```{r}
SCPData_All<-PresAbs %>% filter(gensp %in% "Cottus asper") %>%  group_by(SampleRowID, WaterTemperature, Salinity, Secchi, DO) %>% summarize(Count=sum(Count))
SCPData_All<-as.data.frame(SCPData_All)
```

## Use gbm() from dismo to run model

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

### Setting hyperparameters

```{r}
str(SCPData_All)

Complexity<-4
LearningRate<-0.005
Bag<-0.6  

```

### gbm.step
```{r}
set.seed(124)
system.time( scp_brt_allcount <- dismo::gbm.step(data=SCPData_All, 
                         gbm.x= c(2,3,4,5), # environmental variables.
                         gbm.y= 6, # Binary response variable
                         family = "poisson", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(scp_brt_allcount)

mean(scp_brt_allcount$residuals^2) #mean residual deviance 0.309
#SCPbrt_Age0$
names(scp_brt_allcount) #tells us the components of output
CV_deviance_explained <- (((scp_brt_allcount$self.statistics$mean.null- scp_brt_allcount$cv.statistics$deviance.mean)/scp_brt_allcount$self.statistics$mean.null)*100) 

CV_Correlation<-scp_brt_allcount$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)


 # reponse curves
dismo::gbm.plot(scp_brt_allcount, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.

gbm::plot.gbm(scp_brt_allcount, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(scp_brt_allcount)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(scp_brt_allcount)
find.int#Salinity and DO, Salinity and Water Temp.
scp_brt_allcount$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

scp_brt_allcount$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(scp_brt_allcount,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(scp_brt_allcount,1, 4) #salinity and DO
gbm.perspec(scp_brt_allcount,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(scp_brt_allcount, here("data", "AllSCP_count_brt_gbm_step.rds"))
```


## Data filtering for above 0 counts
```{r}
SCPData_Positive<-PresAbs %>% filter(gensp %in% "Cottus asper", Binary == 1) %>% group_by(SampleRowID, WaterTemperature, Salinity, Secchi, DO) %>% summarize(Count=sum(Count))
SCPData_Positive<-as.data.frame(SCPData_Positive)
summary(SCPData_Positive)
```


## Step 2: Use gbm() from dismo to run model and create outputs

Thanks Nima and Elith! Follow <https://bradleyboehmke.github.io/HOML/gbm.html> for tuning strategy?

### Setting hyperparameters

```{r}
str(SCPData_Positive)

set.seed(124) # for reproducibility 
Complexity<-4
LearningRate<-0.001
Bag<-0.6  

```

### gbm.step
```{r}
system.time( brt3 <- dismo::gbm.step(data=SCPData_Positive, 
                         gbm.x= c(2,3,4,5), # environmental variables.
                         gbm.y= 6, # Binary response variable
                         family = "gaussian", # for counts this would be "poisson"
                         tree.complexity = Complexity, # complexity of the interactions that the model will fit. 5-way interaction since these may all interact with one another? Typical values range from a depth of 3–8 but it is not uncommon to see a tree depth of 1 (J. Friedman, Hastie, and Tibshirani 2001). Smaller depth trees such as decision stumps are computationally efficient (but require more trees); however, higher depth trees allow the algorithm to capture unique interactions but also increase the risk of over-fitting. Note that larger  n  or  p training data sets are more tolerable to deeper trees.
                         learning.rate = LearningRate,  # optimized to end up with >1000 trees. Smaller values make the model robust to the specific characteristics of each individual tree, thus allowing it to generalize well. Smaller values also make it easier to stop prior to overfitting; however, they increase the risk of not reaching the optimum with a fixed number of trees and are more computationally demanding
                         bag.fraction = Bag # default for Elith, amount of input data used each time. the fraction of the training set observations randomly selected to propose the next tree in the expansion. This introduces randomnesses into the model fit. If bag.fraction < 1 then running the same model twice will result in similar but different fits. gbm uses the R random number generator so set.seed can ensure that the model can be reconstructed. 0.5 is ridgeway's preference.
                         ))

# make sure to save the model!
summary(brt3)

mean(brt3$residuals^2) #mean residual deviance 0.309
#SCPbrt_Age0$
names(brt3) #tells us the components of output
CV_deviance_explained <- (((brt3$self.statistics$mean.null- brt3$cv.statistics$deviance.mean)/brt3$self.statistics$mean.null)*100) 

CV_Correlation<-brt3$cv.statistics$correlation.mean

# To pull out one component of the list, use a number (angaus.tc5.lr01[[29]]) or name (angaus.tc5.lr01$cv.statistics); but be careful - some are as big as the dataset! e.g. there will be 1000 fitted values - find this by typing length(angaus.tc5.lr01$fitted)


 # reponse curves
dismo::gbm.plot(brt3, n.plots = 4, write.title= F, rug = T, smooth = TRUE, plot.layout=c(2,2), common.scale = T, y.label = "Fitted Value") #Makes the partial dependence plots.

gbm::plot.gbm(brt3, n.plots=4, return.grid=TRUE)

dismo::gbm.plot.fits(brt3)
#I assume these wtm's are weighted means and are the means used when calculating partial dependence? So each dot is a prediction made from a single tree?


# This code assesses the extent to which pairwise interactions exist in the data.

find.int <- gbm.interactions(brt3)
find.int#Salinity and DO, Salinity and Water Temp.
brt3$self.statistics$correlation[[1]] #The relevant set of evaluation statistics, calculated on the fitted values - i.e. this is only interesting in so far as it demonstrates "evaluation" (i.e. t) on the training data. It should NOT be reported as the model predictive performance.
# The returned object, here named test.int, is a list. The first 2 components summarise the results, first as a ranked list of the 5 most important pairwise interactions, and the second tabulating all pairwise interactions. The variable index numbers in $rank.list can be used for plotting.

brt3$cv.statistics #These are the most appropriate evaluation statistics. We calculate each statistic within each fold (at the identifed optimal number of trees that is calculated on the mean change in predictive deviance over all folds), then present here the mean and standard error of those fold-based statistics.

# You can plot pairwise interactions like this:
par(mfrow=c(1,2))
gbm.perspec(brt3,2, 4) #3D PLOT of Water Temp and Salinity
gbm.perspec(brt3,1, 4) #salinity and DO
gbm.perspec(brt3,1, 2)
#5 is tow duration. 4 is secchi. 1 is water temp.2 is salinity. 3 is DO
# save the model
saveRDS(brt3, here("data", "SCP_positivecount_brt_gbm_step.rds"))
```

## Step 3: Validation

### Splitting out training and testing data.
https://towardsdatascience.com/understanding-gradient-boosting-machines-9be756fe76ab
```{r}
set.seed(124)
inTrain<-createDataPartition(y=SCPData_Positive$Count, p=0.8, list=FALSE) #data is already limited so I decided to use 80% of data to train and will test on 20%. 
SCPData_Positive_Train<-SCPData_Positive[inTrain,]
SCPData_Positive_Test<-SCPData_Positive[-inTrain,]
```

### Running model again

```{r}
LearningRate<-0.05 #had to shift learning rate since only model 80% of the data.
brt4 <- dismo::gbm.step(data=SCPData_Positive_Train, 
                         gbm.x= c(5,6,7,9), # environmental variables. 2 = Water Temperature, 5 = salinity, 3 = DO, 4 = Secchi
                         gbm.y= 22, 
                         family = "poisson", 
                         tree.complexity = Complexity,
                         learning.rate = LearningRate, 
                         bag.fraction = Bag 
                         ) #Learning rate 0.008 gave 650 trees. 0.005 gave 1000 trees
Preds<-predict(object = brtest, newdata = SCPData_Positive_Test, type = "response")

#PredGrid<-gbm.predict.grids(brtest, SCPData_Positive_Test, want.grids=F, n.trees=brtest$gbm.call$best.trees, type="response")
PredBinaries<-as.factor(ifelse(Preds>0.5,1,0))

SCPData_Positive_Test$Binary<-as.factor(SCPData_Positive_Test$Binary)

confusionMatrix(PredBinaries, SCPData_Positive_Test$Binary)

SCPData_Positive_Test$Binary<-as.numeric(SCPData_Positive_Test$Binary)

calc.deviance(SCPData_Positive_Test$Binary, Preds, calc.mean=T, family="bernoulli") #6.890836
library(gbm)

roc(SCPData_Positive_Test$Binary, Preds) #Area under the curve: 0.8112

ROC<-pROC::roc(PredBinaries, SCPData_Positive_Test$Binary)
auc<-round(auc(PredBinaries, SCPData_Positive_Test$Binary),4)
ggroc(ROC,colour = 'steelblue', size = 2) + #create ROC plot
  ggtitle(paste0('ROC Curve', '(AUC= ',auc,') '))+
  theme_minimal()
PredBinaries<-as.numeric(PredBinaries)
cor(PredBinaries, SCPData_Positive_Test$Binary)
plot(PredBinaries, SCPData_Positive_Test$Binary)
```
Roc curve guide: https://www.statology.org/roc-curve-ggplot2/

### 10-fold and loo
Below are 2 validation schemes that we typically use to assess how well the model is predicting. The first is a K-fold cross-validation where the data is split in k number of equal sections (i.e. folds). A model is the trained on one fold and then its predicts to the rest of the folds. It then iterates this process through all the folds. Most studies do 10 folds as this has shown to be a good number. Too low or too high has shown to potentially bias / cause variability in results.

Performance metrics that are calculated are deviance explained, AUC, and TSS. Deviance explained is a metric for how much variance did our model explain (you can think of is as R squared) and AUC and TSS are assessing predictive skill. Since you are using abundance data you may need to change that to RMSE or MAE (I’d suggest taking a look at the Metrics package for that)

NOTE: whether you are using gbm.fixed or gbm.step to fit your BRTS, you will have to make a slight change to this function below. I’ve commented lines out since I made the model above using gbm.fixed but if used gbm.step I would used the code on those lines instead.

```{r}
#k fold cross validation function

eval_kfold_brt <- function(dataInput, gbm.x, gbm.y, learning.rate = LearningRate, k_folds = 10, tree.complexity = Complexity, bag.fraction = Bag){
  dataInput$Kset <- dismo::kfold(dataInput, k_folds) #randomly allocate k groups
  Evaluations_kfold_BRT <- as.data.frame(matrix(data=0,nrow=5,ncol=4)) 
  colnames(Evaluations_kfold_BRT) <- c("k","Deviance","AUC","TSS")
  counter=1
  for (k in 1:k_folds){
    print(k)
    train <- dataInput[dataInput$Kset!=k,]
    test <- dataInput[dataInput$Kset==k,]
    
     brt.k <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                             gbm.y = gbm.y, 
                              family="poisson", 
                             tree.complexity = tree.complexity,
                             learning.rate = learning.rate, 
                             bag.fraction = bag.fraction)
    
    preds <- gbm::predict.gbm(brt.k, test,
                              n.trees=brt.k$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.k)
    d <- cbind(test[gbm.y], preds)
    pres <- as.numeric(d[d[,1]==1,2])
    abs <- as.numeric(d[d[,1]==0,2])
    e <- dismo::evaluate(p=pres, a=abs)
    Evaluations_kfold_BRT[counter,1] <- k
    Evaluations_kfold_BRT[counter,2] <- dev
    Evaluations_kfold_BRT[counter,3] <- e@auc
    Evaluations_kfold_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    counter=counter+1 
  }
  return(Evaluations_kfold_BRT)
}
  
  
brt_SDM_k <- eval_kfold_brt(dataInput = LongfinSmeltYOYData_Positive,
                        gbm.x = c(2,3,4,5),
                        gbm.y=6, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity,
                        k_folds = 10)
brt_SDM_k
```

The other validation scheme that we do is a Leave one [year] out cross-validation. Just like 5 folds but now were trained the model on all the data but one year. And then we predict to that one year that was not included. This give us a sense out how well can we predict year by year as there may be some environmental variability over time.

NOTE: Same as above of changing the code whether you are using gbm.fixed or gbm.step

```{r}
#leave one year out function

eval_loo_brt <- function(pres, abs, gbm.x, gbm.y, learning.rate = LearningRate, tree.complexity = Complexity, bag.fraction = Bag){
  
  pres$year <- pres$Year
  abs$year <- abs$Year
  
  ## setup output df
  Evaluations_LOO_BRT <- as.data.frame(matrix(data = 0, nrow = 1, ncol = 5))
  colnames(Evaluations_LOO_BRT) <- c("k","Deviance","AUC","TSS","n_pres")
  counter=1
  
  u_years <- unique(pres$year)
  for (y in u_years){
    
    print(paste('Running ', which(u_years == y), ' of ', length(u_years), sep=''), sep='')
    
    pres_train <- pres[which(pres$year != y),]
    ## absences are sampled from the full input absence df
    abs_train <- abs[sample(which(abs$year != y), size = nrow(pres_train), replace=F),]
    train <- rbind(pres_train, abs_train)
    
    pres_test <- pres[which(pres$year == y),]
    abs_test <- abs[sample(which(abs$year == y), size = nrow(pres_test), replace=F),]
    test <- rbind(pres_test, abs_test)
    
    
     brt.loo <- dismo::gbm.step(data=train, gbm.x= gbm.x, 
                                gbm.y = gbm.y, 
                                family="poisson", 
                                tree.complexity = tree.complexity,
                               learning.rate = learning.rate, 
                               bag.fraction = bag.fraction)
    
    ## make predictions for eval
    preds <- gbm::predict.gbm(brt.loo, test,
                              n.trees=brt.loo$gbm.call$best.trees, 
                              type="response")
    
    dev_eval3<-function(x){
      null <- x$self.statistics$mean.null #use with gbm.step
      res <- x$self.statistics$mean.resid #use with gbm.step
      dev=((null - res)/null)*100
      return(dev)
    }
    dev<-dev_eval3(brt.loo)
    d <- cbind(test[,gbm.y], preds)
    pres_y <- as.numeric(d[d[,1] == 1,2])
    abs_y <- as.numeric(d[d[,1] == 0,2])
    e <- dismo::evaluate(p = pres_y, a = abs_y)
    
    Evaluations_LOO_BRT[counter,1] <- y
    Evaluations_LOO_BRT[counter,2] <- dev
    Evaluations_LOO_BRT[counter,3] <- e@auc
    Evaluations_LOO_BRT[counter,4] <- max(e@TPR + e@TNR-1)
    Evaluations_LOO_BRT[counter,5] <- length(which(train[,gbm.y] == 1))
    
    counter=counter+1 
  }
  return(Evaluations_LOO_BRT)
}

brt_SDM_loo <- eval_loo_brt(pres = LongfinSmeltYOYData_Positive[which(LongfinSmeltYOYData_Positive$Binary == 1),],
                        abs = LongfinSmeltYOYData_Positive[which(LongfinSmeltYOYData_Positive$Binary == 0),],
                        gbm.x = c(2,3,4,5),
                        gbm.y=6, learning.rate = LearningRate,
                        bag.fraction = Bag, tree.complexity = Complexity)
brt_SDM_loo

#combining the model output, and both validation results into a list to save
eval <- list(brt = SCPbrt_Age0, brt_k = brt_SDM_k, brt_loo = brt_SDM_loo)

eval
#now save it

saveRDS(eval, here("data","YOYLongfinSmelt_brt_SDM_eval_gbm_step.rds"))
```


## Step 4: Analyzing Response Curves & Variable Importance

Look at Nima's code for pretty plots: <https://github.com/nfarchadi/heatwave_impacts_on_fisheries/blob/main/scripts/4_plots/FS3_FS4_response_relimport_plots.R>

### Nima plots
```{r}
# Fig. S3. The relative importance of drivers in distribution models
# Fig. S4. Response curves of each environmental variable from distribution models


###----response plots----####

brt_LongfinSmelt <- here("data","LongfinSmelt_brt_SDM_eval_gbm_step.rds") %>% readRDS()


# get the matrix out of the `plot.gbm`
var<-brt_LongfinSmelt$brt$var.names
response_data<-data.frame()

for (i in 1:length(var)){
response_plot_data <- gbm::plot.gbm(brt_LongfinSmelt$brt,
                                 i.var = var[i],
                                 return.grid = TRUE)

response_plot_data<-response_plot_data %>% gather("variable","x",-y)

response_data<-rbind(response_data, response_plot_data)

}

#Get it on the same scale as elith
#response_data$y <- response_data[,2] - mean(response_data[,2])
#response_data$y<-1/(1+exp(-response_data$y))

#response_data$y<-scale(response_data$y, scale = FALSE)

#Add a smoothed line to an active plot
#loess(y ~ x, span = 0.3)

response_data %>%
  ggplot() + 
  geom_line(aes(x=x, y=y), color = "lightseagreen", size =1) + 
  facet_wrap(~variable, scales = "free", nrow = 5) +
  labs(x = "Variable Values",
       y = "Marginal Effect")+
  theme_bw() + 
  theme(panel.spacing = unit(.30, "lines"),
        strip.text = element_text(size=10),
        strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0,0,.05,0, "cm")))
  
ggsave(here("Plots","LongfinSmelt_response_plots.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","LongfinSmelt_response_plots.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)

###----variable contribution----####
rel.inf<-data.frame(Variable = brt_LongfinSmelt$brt$contributions$var,
                        rel.inf=brt_LongfinSmelt$brt$contributions$rel.inf)# %>% 



rel.inf %>% 
  ggplot()+
  geom_bar(aes(x = reorder(Variable, -rel.inf),y=rel.inf), fill = "lightseagreen",
           stat="identity", position=position_dodge())+
  theme_bw() +
  labs(x = "Environmental Variable",
       y = "Relative Importance (%)")+
  theme(axis.text.x = element_text(angle = 45,hjust=1))

ggsave(here("Plots","LongfinSmelt_relative_importance.png"),
       width = 7, height = 6, units = "in", dpi = 300)
#ggsave(here("Plots","FS3_relative_importance.svg"),
#       width = 7, height = 6, units = "in", dpi = 300)
```

When looking at partial dependence plots, the y-axis represents the marginal impact of the independent variable to the dependent variable. If at 0, then there is 0 impact...


### Looking at caret outputs for optimal model (ROC and Kappa)

```{r}
gbmGrid <- expand.grid(interaction.depth = Complexity, #tree complexity
                       n.trees = 1100,
                       shrinkage = LearningRate,
                       n.minobsinnode = 10 )

gbmGrid %>% head() #take a look at what we just made

#specify type of resampling
fitControlKappa <- trainControl(## 10-fold CV
                           method = "repeatedcv", #"cv" if roc metric is used below.
                           number = 10, #Can increase! 
                           repeats=3, #Delete for roc
                           )

#specify type of resampling
fitControlROC <- trainControl(## 10-fold CV
                           method = "cv", 
                           number = 10, #Can increase! 
                           classProbs = TRUE, ## Estimate class probabilities for classification models if ROC
                           summaryFunction = twoClassSummary #a function to compute performance metrics across resamples. 
                           )

set.seed(124) #setting the seed for reproducibility. BRT have randomness in them so if you want to make sure you are getting the same result each time make sure to set the seed

gbmFitKappa <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = SCPData_Positive, 
                method = "gbm", 
                trControl = fitControlKappa, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "Kappa" #Or ROC 
                )

gbmFitKappa

gbmFitROC <- train(PresAbs ~ WaterTemperature + Secchi + DO + Salinity, #caret wants non-numeric presence, absence. Might need to change it to Yes and No.
                data = SCPData_Positive, 
                method = "gbm", 
                trControl = fitControlROC, 
                bag.fraction=Bag,
                na.action=na.pass,
                distribution="bernoulli",
                verbose = FALSE, #for gbm's this is needed.
                ## Now specify the exact models to evaluate:
                tuneGrid = gbmGrid, #what we made above
                metric = "ROC"  
                )
gbmFitROC
# save the model
#saveRDS(gbmFit, here("data", "brt_gbm_fit.rds"))

# we can plot the results too which will be easier to interpret
ggplot(gbmFit)
```

Stochastic Gradient Boosting 3505 samples   4 predictor   2 classes: 'absent', 'present' No pre-processingResampling: Cross-Validated (10 fold, repeated 3 times) Summary of sample sizes: 3155, 3154, 3155, 3154, 3155, 3155, ... Resampling results:  Accuracy  Kappa      0.932194  0.1785341Tuning parameter 'n.trees' was held constant at a value of a value of 0.08Tuning parameter 'n.minobsinnode' was held constant at a value of 10

Stochastic Gradient Boosting 

3505 samples
   4 predictor
   2 classes: 'absent', 'present' 

No pre-processing
Resampling: Cross-Validated (10 fold) 
Summary of sample sizes: 3154, 3155, 3155, 3155, 3155, 3154, ... 
Resampling results:

  ROC       Sens       Spec     
  0.774977  0.9838776  0.1422078

Tuning parameter 'n.trees' was held constant at a value of
 a value of 0.08
Tuning parameter 'n.minobsinnode' was
 held constant at a value of 10
 
# ALL AGES TOGETHER 1 row for each sampling event for
```{r}
PricklySculpinDataPA_AllAges<-PricklySculpinData %>% dplyr::group_by(SampleRowID, WaterTemperature, DO, Secchi, Salinity, Year, Month, StationCode) %>% dplyr::summarize(Count=sum(Count)) %>% mutate(PresAbs = if_else(Count > 0, "present", "absent"), Binary=if_else(Count > 0, 1, 0)) #For all ages together.

PricklySculpinDataPA_AllAges$PresAbs<-as.factor(PricklySculpinDataPA_AllAges$PresAbs)

PricklySculpinDataPA_AllAges$Year<-as.factor(PricklySculpinDataPA_AllAges$Year)

PricklySculpinDataPA_AllAges$Month<-as.factor(PricklySculpinDataPA_AllAges$Month)

PricklySculpinDataPA_AllAges$StationCode<-as.factor(PricklySculpinDataPA_AllAges$StationCode)

PricklySculpinDataPA_AllAges<-as.data.frame(PricklySculpinDataPA_AllAges)
```

### Summaries of catch
```{r}

print(PricklySculpinDataPA_AllAges %>% group_by(Year) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100), n=26) #Boom and bust years, not great for loo validation.

PricklySculpinDataPA_AllAges %>% group_by(StationCode) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) %>% print(n=24) #Some are heavily favored.

PricklySculpinDataPA_AllAges %>% group_by(Month) %>% dplyr::summarize(N=n(), Count = sum(Count), BinarySum = sum(Binary), Probability=(BinarySum/N)*100) #Never in January and February

#Overall capture probability
OverallProb<-(sum(PricklySculpinDataPA_AllAges$Binary)/length(unique(PricklySculpinDataPA_AllAges$SampleRowID)))*100
```

`OverallProb`% probability of capturing all ages PricklySculpin.


# NOTES

-   If number of trees below 1k, need to lower learning rate.
-   There may be a difference in the caret output and gbm output (5%). To avoid this can use the default grid to optimize parameters and use predict to have the same results... Or don't worry about it.
