---
title: "SMFS_WranglingForThesis"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
Summary:
Take the output from SMFS_Wrangling and wrangle a bit for thesis modeling purposes (presence absence data). 

1. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data
```{r}
#Wrangling SMFS 
library(tidyverse)
library(rlang)
library(readr)
library(lubridate)
library(lfstat)
library(readxl)

WrangledSMFSForIntegration<-read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_WithAbsences_062424.csv")
glimpse(WrangledSMFSForIntegration)
```

# Filtering
 Removing NA's even though BRTs can handle them because GAMs cant (I think).
```{r}

#Now I want to constrain the data to the years of interest (2011 to now)
str(WrangledSMFSForIntegration)
WrangledSMFSForIntegration$WaterYear<-as.character(WrangledSMFSForIntegration$WaterYear)#when converting the factor to a number, need to first make it a character and THEN numeric. Otherwise, the number gets changedfor some reason.
WrangledSMFSForIntegration$WaterYear<-as.numeric(WrangledSMFSForIntegration$WaterYear) #need to make water year numeric so I can filter out years easily.
WrangledSMFSForIntegration$Year<-as.numeric(WrangledSMFSForIntegration$Year) #need to make water year numeric so I can filter out years easily.

NumberSamples1<-unique(WrangledSMFSForIntegration$SampleRowID) #3801. This is the number of unique samples.
NumberTows<-unique(WrangledSMFSForIntegration$TrawlRowID) #Same
#Removing NAs
#NOTE: There is info for other wq variables. So I am losing some amount of good info here... 
SMFS_NoNA<-WrangledSMFSForIntegration %>% filter(!TowDuration %in% NA, !DO %in% NA, !Salinity %in% NA, !Secchi %in% NA, !WaterTemperature %in% NA, !TideCode %in% NA, !StationCode %in% NA) #Remove NA rows for Tow Duration, DO, Salinity, Secchi, and Water Temp (Important WQ). Also removed NA's for other variables of interest for good measure. 

SMFS_Thesis<-SMFS_NoNA %>% filter(!StationCode%in% c("BL", "CHA1", "CHA2", "COGR", "CRD1", "CRD2", "CRD3", "CRD3.5", "CRD4", "FH1", "GR3", "GR4", "HL2", "IBI1", "LU1", "PT3", "SD2", "SUVOL", "TRDB2", "MWB1", "MWB2", "MWS1", "MWS2", "MZN3", "SB0_5"))

which(is.na(SMFS_Thesis$TowDuration)) #Tow Duration NAs removed.
NumberSamples2<-unique(SMFS_NoNA$SampleRowID) #3760

#Which sites to keep?
#CODE NOT RELEVANT TO DATA WITH ABSENCES AS IT INFLATES SAMPLES. SMFS_Thesis %>% ggplot() + geom_bar(aes(StationCode))+geom_hline(aes(yintercept=1000), color="red") +theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 #adding a cutoff of 1000 to be included in the study.

NumberSamples3<-unique(SMFS_Thesis$SampleRowID) #3505 reduced to 3433 after removing NAs and undersampled locations. This is the number of unique samples. Expect 3433 samples * 6 species to yield 20,598 rows to have fish shown for each sample.

```


## Selecting species of interest
```{r}
#Do I have instances of catching only large fish not well sampled?
Inquiry<- SMFS_Thesis %>% filter(gensp %in% c("Alosa sapidissima", "Spirinchus thaleichthys", "Cottus asper", "Gasterosteus aculeatus", "Menidia beryllina", "Tridentiger barbatus")) #7479
species<-c("Alosa sapidissima", "Spirinchus thaleichthys", "Cottus asper", "Gasterosteus aculeatus", "Menidia beryllina", "Tridentiger barbatus")

AgeTable<-Inquiry %>% group_by(gensp, AgeClass) %>% dplyr::summarise(CountSum = sum(Count))
```

For presence absence models, I can get rid of plus counts AND I can remove large fish.
Based on plots of wq responses by age class, large fish are as folows:
LFS Age 1+
ASH Age-1+
ISS Age-1
SCP Age-1+
SKG Age-1+
STBK have no age classes.


# Add presence/absence info
```{r}
presabs<-Inquiry %>% mutate(PresAbs = if_else(Count > 0, "present", "absent"), Binary=if_else(Count > 0, 1, 0))

WieLang<-unique(presabs$SampleRowID) #3505
```

Plus counts unaltered but may want to update them so that I can create a filter to remove non-targeted fish sizes.

Mikaela: For presence/absence model, we don't care about plus counts. Remove all plus count data. Filter on tows that measured big fish. Remove those fish/observations. And then for all tows, label each tow as 'presence' or 'absent'. If there was a tow that only measured big fish, it will look like no fish were captured because you removed all instances of big fish measurements. This avoids dealing with plus count, which we don't need for the presence/absent model. I



# Save
```{r}
#Save New DF
write.csv(presabs, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_Thesis_PresAbs_062524.csv")

```