---
title: "AgeClass"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---
```{r}
library(fuzzyjoin)
library(tidyverse)
library(rlang)
library(cowplot)
library(readr)
SMFS_05202023 <- read_csv("Data/SMFS_05202023.csv")
str(SMFS_05202023)
library(lubridate)
SMFS_05202023$Year<-year(SMFS_05202023$SampleDate)
SMFS_05202023$Month<-month(SMFS_05202023$SampleDate)
library(lfstat)
SMFS_05202023$WaterYear<-water_year(SMFS_05202023$SampleDate, as.POSIX = FALSE, origin="usgs")

#Now I just want otter trawl data
SMFS_OTR_05202023<-SMFS_05202023 %>% filter(MethodCode %in% "OTR") #filters otter trawl samples from table

#Now I want to constrain the data to the years of interest (2011 to now)
str(SMFS_OTR_05202023)
SMFS_OTR_05202023$WaterYear<-as.character(SMFS_OTR_05202023$WaterYear)#when converting the factor to a number, need to first make it a character and THEN numeric. Otherwise, the number gets changedfor some reason.
SMFS_OTR_05202023$WaterYear<-as.numeric(SMFS_OTR_05202023$WaterYear) #need to make water year numeric so I can filter out years easily.
SMFS_OTR_Thesis<-SMFS_OTR_05202023 %>% filter(WaterYear > 2010) #Filtering out water years which overlap with South Bay Otter Trawl Survey for my thesis work.


library(readxl)
SMFS_AgesBySizeMo <- read_excel("Data/SMFS_AgesBySizeMo.xlsx")
#The N/A class throws off r. I need to rename that value 
SMFS_AgesBySizeMo$Class[SMFS_AgesBySizeMo$Class == 'N/A'] <- 'NA_Class'

FishOfInterest<- c("SB", "ASH", "TFS", "ISS", "SCP", "SF", "TP", "LFS", "YFG")

SMFS_OTR_Thesis<-SMFS_OTR_Thesis %>% filter(OrganismCode %in% FishOfInterest)

# Perform a conditional left join based on range conditions
merged_data <- fuzzy_left_join(SMFS_OTR_Thesis, SMFS_AgesBySizeMo, 
                               by = c("OrganismCode" = "OrganismCode",
                                      "StandardLength" = "Min", "StandardLength" = "Max", "Month"),
                               match_fun = list(`==`, `>=`, `<=`, `==`))
#It worked! 

# Select the desired columns and remove duplicates
#merged_data <- merged_data %>%
 # select(-Month.x, -Min.x, -Max.x) %>%
  #rename(Class = Class.y)
# Remove duplicate Month column from SMFS_AgesBySizeMo
merged_data <- merged_data %>% rename(Month=Month.x)
merged_data <- merged_data %>% rename(OrganismCode=OrganismCode.x)

# Check the resulting dataframe
#glimpse(merged_data)
#Okay, before adding 0's I will fix the plus catch data.

unique(merged_data$CatchComments) #Getting all of the catch comments that I may have to deal with. 
#Lets check that these comments only apply to fish with no length data measured, otherwise I'll have to do more complicated wrangling...

#check<-merged_data %>% filter(StandardLength > 0) 
#unique(check$CatchComments) #Comments are NA, "estimated" and "had eaten SCP and EXO"... So not going to be an issue.

#Alright, now I will tell r that any time "yoy", "YOY", or "young of year" pop up, that we will replace NA with "Age-0" in the Class column.

glimpse(merged_data)

# Iterate through the rows of the dataset and update "Class" based on "CatchComments"
for (i in 1:nrow(merged_data)) { #loop through each row of dataset
  comment <- merged_data$CatchComments[i] #Check if comments containe yoy variations
  if (!is.na(comment) && grepl("(?i)yoy|young of year", comment)) {
    merged_data$Class[i] <- "Age-0" #if it does, update class column
  }
}

# Print the updated dataset
print(merged_data)

# Replace values in the "Class" column based on "CatchComments"
updated_data <- merged_data %>%
  mutate(Class = if_else(
    !is.na(CatchComments) & grepl("(?i)yoy|young of year", CatchComments),
    "Age-0",
    Class
  )) #almost worked. Seems to have been arbitrarily added "Age-0" all over the place... NEED TO DEBUG!

# Print the updated dataset
print(updated_data)

# Step 1: Filter rows where Class is not "NA_Class" and Counted is greater than 0
df_filtered <- merged_data %>%
  filter(!is.na(Class), Class != "NA_Class", Counted > 0)

# Step 2: Calculate the total count for each SampleRowID and OrganismCode
df_total_counts <- df_filtered %>%
  group_by(SampleRowID, OrganismCode) %>%
  summarize(TotalCount = sum(Counted), .groups = "drop")

# Step 3: Calculate the proportion for each age class
df_proportions <- df_Class_NA %>%
  left_join(df_total_counts, by = c("SampleRowID", "OrganismCode")) %>%
  mutate(
    Proportion = ifelse(
      Class != "NA_Class",
      Counted / TotalCount,
      0
    )
  )

# Step 4: Calculate the total "NA" counts for each SampleRowID
df_na_counts <- df_proportions %>%
  filter(is.na(Class)) %>%
  group_by(SampleRowID) %>%
  summarize(NA_Count = sum(Counted), .groups = "drop")

# Step 5: Distribute the "NA" counts proportionally to other age classes
df_final <- df_proportions %>%
  left_join(df_na_counts, by = "SampleRowID") %>%
  mutate(
    Counted = ifelse(
      !is.na(Class),
      round(Counted + Proportion * NA_Count),
      0
    )
  ) %>%
  select(-TotalCount, -Proportion, -NA_Count)

#Now add zeros!
SMFS_OTR_Thesis_AgeClassWithZeros<-merged_data %>% 
  complete(nesting(SampleRowID, 
                   WaterTemperature, 
                   Secchi, 
                   DO, 
                   Salinity, 
                   MethodCode, 
                   StationCode, 
                   SampleDate, 
                   SampleTime, 
                   PctSaturation, 
                   SpecificConductance, 
                   TideCode, 
                   ElecCond, 
                   TowDuration, 
                   TrawlComments, 
                   Year,
                   Month,
                   WaterYear), 
           OrganismCode, Class,
           fill=list(num=0)) %>% as.data.frame() #Didn't add 0 values...Also ended up with more rows than expected? I have more rows than expected because there are multiple SIZES which each get their own row as well. So if I care about size/age classes, will have to bin those first and then add zeros. For now, move forward here.

SMFS_OTR_Thesis_AgeClassWithZeros$Count[is.na(SMFS_OTR_Thesis_AgeClassWithZeros$Count)] <- 0
#NOW it has zeros
```

# Now Plot For each age class for each species of interest
```{r}
#First, what are the age classes for each species?
ClassCountTable<-SMFS_OTR_Thesis_AgeClassWithZeros %>% group_by(OrganismCode, Class) %>% summarise(Total=sum(Count))
# Filter rows with the specified 'FishOfInterest'
filtered_data <- SMFS_OTR_Thesis_AgeClassWithZeros %>%
  filter(OrganismCode %in% FishOfInterest)

# Count the number of distinct 'SampleRowID' values in the filtered data
n_distinct(filtered_data$SampleRowID)

```

ASH: 1027 Age 0, 0 Age 1, 296 Age 1+, 0 Age 2+
ISS: no age class info: 929 + 328
SB: 10534 Age 0, 4069 Age 1, 0 Age 1+, 975 Age 2+, 5082 NA
Shoot, there are a bunch of plus counted fish in here.
I want to incorporate them. 

#Lets see how many fish have no standard length associated with them.
```{r}
NoSL<-SMFS_OTR_Thesis %>% filter(StandardLength == 0, OrganismCode %in% FishOfInterest)
n_distinct(NoSL$TrawlRowID)
TotalNoSL<-NoSL %>% summarise(Total=sum(Count)) #Total number of plus counted fish in dataset = 6984
WithSL<-SMFS_OTR_Thesis %>% filter(!StandardLength == 0, OrganismCode %in% FishOfInterest) %>% summarise(Total=sum(Count)) #Total number of fish with length data = 34641
#Total number of fish in dataset = 41,625
#16% of the data is plus count!!!
```

Be careful to not assign lengths to fish that were plus counted because they escaped. There should be some sort of rule where we only assign sizes if there were over 30 of a species in the trawl.
Also, some of the entries have comments indicating the size of the fishes (though aren't always proportionally calcuable)
Let's grab the Trawl ID's from these and then pull all of the associated trawls out into one table.

```{r}
#start with a subset for now
PlusCountSamples<-merge(NoSL, SMFS_OTR_Thesis, by = c("SampleRowID", "OrganismCode")) #merging the thesis data before adding zeros with the df I made of plus count fish to retain all records of relevant fish (organism code) from the same trawls as the plus counted fish.
n_distinct(PlusCountSamples$SampleRowID) #verifying that I still have the same number of trawls in consideration.
SampleSum<-PlusCountSamples %>% filter(!StandardLength.y==0) %>%  group_by(SampleRowID, OrganismCode) %>% summarise(Total=sum(Count.y)) #Taking the plus count sample tows and summing the number of counts of fish for each species to see how many fish were measured before plus counting
SampleOver30<-SampleSum %>% filter(Total>=30) #filtering out the tows which reached the minimum number of recorded lengths 
#make a list of sample row IDs for these 
SampleOver30List<-SampleOver30$SampleRowID
length(SampleOver30List)

PlusCountSamplesOver30<-PlusCountSamples %>% filter(SampleRowID %in% SampleOver30List)

SMFS_OTR_ThesisFish<- SMFS_OTR_Thesis %>% filter(OrganismCode %in% FishOfInterest)

#I want to create a column indicating that a fish with 0 SL is a plus counted fish that we want to estimate the size of (had over 30 counts). I can use catch row ID to do this!


# Step 1: Calculate the total standard length for each SampleRowID
df_total_Class <- filtered_data %>% #has the data for just fish of interest
  filter(SampleRowID %in% SampleOver30List) %>% #
  group_by(SampleRowID, OrganismCode, Class) %>% 
  summarise(Counted=sum(Count)) 


df_Class_NA<-df_total_Class %>%
  group_by(SampleRowID, OrganismCode)%>%
  filter(OrganismCode %in% c("SB", "SCP", "TP", "LFS"),sum(Counted) >= 30)
print(df_Class_NA, n=100)
n_distinct(df_Class_NA$SampleRowID) #Not sure why there are 3 samples less than what we started with when I was just trying to delete the extra rows...
df_Class_NoNA<-df_total_Class %>%
  group_by(SampleRowID, OrganismCode)%>%
  filter(sum(Counted, na.rm=TRUE) >= 30)
print(df_total_Class, n=100)
```
For each instance that there is "NA" in Class with a count above 0, 
I want to take that count and distribute those counts to the other age classes that do have counts, in proportion to the age classes from that same sample (ie same sample row ID) that have counts. 
For example, Sample Row ID {001EFF22-FE47-4468-92F6-6F1531090921}  has  29 "Age-0" "SB" and 1 "Age 2+" "SB" as well as 6 SB that have no age class value ("NA"). I want to take those 6 "NA" fish and assign them proportionally to the other age classes: ((29/30) * 6) added to "Age-0" and ((1/30) * 6) added to "Age 2+" for a total of 29+5.8 = 34.8 "Age-0" SB for Sample Row ID {001EFF22-FE47-4468-92F6-6F1531090921}  and 1+0.2 = 1.2 "Age 2+" SB and 0 "NA" SB.

```{r}
# Step 1: Filter rows where Class is not "NA_Class" and Counted is greater than 0
df_filtered <- df_Class_NA %>%
  filter(!is.na(Class), Class != "NA_Class", Counted > 0)

# Step 2: Calculate the total count for each SampleRowID and OrganismCode
df_total_counts <- df_filtered %>%
  group_by(SampleRowID, OrganismCode) %>%
  summarize(TotalCount = sum(Counted), .groups = "drop")

# Step 3: Calculate the proportion for each age class
df_proportions <- df_Class_NA %>%
  left_join(df_total_counts, by = c("SampleRowID", "OrganismCode")) %>%
  mutate(
    Proportion = ifelse(
      Class != "NA_Class",
      Counted / TotalCount,
      0
    )
  )

# Step 4: Calculate the total "NA" counts for each SampleRowID
df_na_counts <- df_proportions %>%
  filter(is.na(Class)) %>%
  group_by(SampleRowID) %>%
  summarize(NA_Count = sum(Counted), .groups = "drop")

# Step 5: Distribute the "NA" counts proportionally to other age classes
df_final <- df_proportions %>%
  left_join(df_na_counts, by = "SampleRowID") %>%
  mutate(
    Counted = ifelse(
      !is.na(Class),
      round(Counted + Proportion * NA_Count),
      0
    )
  ) %>%
  select(-TotalCount, -Proportion, -NA_Count)

# View the resulting data frame
View(df_final)
#I think that worked!!!

#Alright, so now I have to either incorporate those values into the original dataframe, replacing by SampleRowID. OR I have to code a way to get this done in the original dataframe...


# Create a new column 'Counted' and set it to 0 for all rows
#df_final <- df_final %>%
 # mutate(Counted = 0)

# Update the 'Counted' column based on conditions
#df_final <- df_final %>%
 # group_by(SampleRowID, OrganismCode) %>%
  #mutate(
   # Counted = case_when(
    #  Class %in% c("Age-0", "Age-1", "Age-1+") & !is.na(Count) ~ sum(Count, na.rm = TRUE),
     # Class %in% c("Age-2+") ~ sum(Count, na.rm = TRUE),
      #TRUE ~ Counted
    #)
  #) %>%
  #ungroup()

# Print the first 100 rows
#print(df_final, n = 100)


```

```{r}
#From chatgpt:

# Step 1: Filter rows where Class is not "NA_Class" and Counted is greater than 0
#df_filtered <- df_total_Class %>%
 # filter(!is.na(Class), Class != "NA_Class", Counted > 0)

# Step 2: Calculate the total count for each SampleRowID and OrganismCode
#df_total_counts <- df_filtered %>%
 # group_by(SampleRowID, OrganismCode) %>%
  #summarize(TotalCount = sum(Counted), .groups = "drop")

# Step 3: For rows with "NA" in Class and Counted greater than 0, distribute the counts proportionally
#df_distributed <- df_filtered %>%
 # left_join(df_total_counts, by = c("SampleRowID", "OrganismCode")) %>%
  #mutate(Proportion = Counted / TotalCount) %>%
  #group_by(SampleRowID, OrganismCode) %>%
  #complete(Class, fill = list(Counted = 0)) %>%
  #mutate(
   # Counted = ifelse(Class == "NA_Class", round(sum(Counted * Proportion)), Counted)
  #) %>%
  #ungroup()

# Step 4: Calculate the updated total count for each SampleRowID and OrganismCode after distributing counts
#df_updated_counts <- df_distributed %>%
 # group_by(SampleRowID, OrganismCode) %>%
  #summarize(UpdatedTotalCount = sum(Counted), .groups = "drop")

# Step 5: Replace the "NA" rows with the updated counts for each age class
#df_final <- df_total_Class %>%
 # left_join(df_updated_counts, by = c("SampleRowID", "OrganismCode")) %>%
  #mutate(
   # Counted = ifelse(is.na(Class), UpdatedTotalCount, Counted),
    #Class = ifelse(is.na(Class), "NA_Class", Class)
  #) %>%
#  select(-UpdatedTotalCount)

# View the resulting data frame
#View(df_final)
#print(df_final, n=50)

```
In this code, we first calculate the total count for each unique combination of SampleRowID and OrganismCode. Then, we merge the total counts back to the original dataframe and calculate the proportion of each age class count to the total count. Finally, we use the ifelse function within a grouped operation to replace "NA" with the corresponding age class based on the calculated proportions.

Make sure to replace 'df' with the actual name of your data frame and adjust the column names as needed to match your data.










