---
title: "SMFS Wrangling"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this rmd, I start with the output from SMFSDataTableMerging (SMFS_09192023) to complete the following objectives:

Create Fully wrangled dataframe by
- Create SMFS_Integrate that has columns for Year, Month, and Water Year, Filters to only keep otter trawls and samples after 2010; Add a column for Organism Code scientific names to match SBOTS (gensp)
- Give each measured fish its own row (OneMeasuredFihOneRow) and separate out all of the non-measured stuff (PlusCountDF)
- I separate out fish from PlusCountDF (PlusCount_OC) and assign lengths to plus count fish based on measured fish of the same tow (PlusAssigned -> PlusJoined to give fish sampling info).
- Put all organisms back in same df ()
- Add age class info
- Compare and fish plus assignments for YOY fish that don't match catch comments. (Other sizes of fish harder to do)
- Add zeros for when fish aren't caught (each species, each size class)
- Save those changes as SMFS_Integrate



*More details on how I expand plus count data:

Poisson: I take each tow, species combination and create a poisson distribution of the sizes of measured fish. I then use that distribution to randomly assign lengths to the plus count fish. From there I assign the plus count fish age classes.

Checking the comments to ground-truth:
I check the outputs of each approach against the comments that detail which fish are YOY (not all YOY recieve comments). Results for poisson show that only 9 of the 3629 fish from the poisson have been misidentified as older than age-0 despite having comments that they are age 0. Those fish have the following plus ID's: 1099, 1109, 1131, 1140, 1144, 1174, 4708, 8225, 8226 

There are some fish with no length assignment because they have no measured fish to be based off of. They were one-offs that length measurements were missed on. Levi wants me to deal with them by basing their lengths on lengths of the day.

2. Investigate NA's in the dataset to evaluate the need to interpolate missing environmental data. I create a dataframe of missing environmental data samples called SMFS_NA_WaterQuality and find that there are 46 sampling events with missing wq data. Of these:
> 8 have NA Tow Duration (not real tows)
> 2 from Boynton in July 2013 missing DO and PctSat (BY1, 3)
> 2 from Cutoff in July 2013 missing DO and PctSat (CO1, 2)
> 4 from Denverton (July '13 DV2 NA DO & PctSat; August '13 DV2 NA Secchi; June '12 DV3 NA DO & PctSat; July '13 DV3 NA DO & PctSat)
> 3 from Good Year missing DO & PctSat from July 2013 (GY1,2,3)
> 1 from HL(2) in May 2018 missing Temp, Salinity, DO, PctSat, SpecificCond 
> 1 from Luco (LU1) July 2013 missing DO & PctSat
> 5 from Montezuma slough (MZ1 July 2013 NA DO & PctSat; MZ1 Jan 2015 NA DO & PctSat; MZ2 July 2013 NA DO & PctSat; MZ2 Jan 2015 NA DO & PctSat; MZ6 July 2013 NA DO & PctSat)
> 2 from MZN 3 (July 2013 NA DO & PctSat; Dec 2013 NA Temp, Salinity, DO, PctSat, SpecificCond)
> 3 from Nurse (NS2 July 2013 NA DO & PctSat; NS3 July 2013 NA DO & PctSat; NS3 August 2014 NA Secchi)
> 5 from Peytonia (PT1 August 2022 NA DO & PctSat; PT1 July 2013 NA DO & PctSat; PT2 May 2018 NA Temp, Salinity, DO, PctSat, SpecificCond; PT2 July 2013 NA DO & PctSat; PT2 August 2022NA DO & PctSat)
> 3 from SB (SB1 July '13 NA DO & PctSat; SB2 May 2018 NA Temp, Salinity, DO, PctSat, SpecCond; SB2 July 2013 NA DO & PctSat)
> 1 from Shelldrake (SD2) May 2018 NA Temp, Salinity, DO, PctSat, SpecificCond
> 6 from Suisun (SU1 July 2013 NA DO & PctSat; SU2 July 2013 NA Secchi, DO, PctSat; SU2 August 2013 NA salinity & SpecCond; SU3 July 2013 NA DO & PctSat; SU3 Oct 2013 NA DO & PctSat; SU4 July 2013 NA DO & PctSat)

# Loading Data
```{r}
#Wrangling SMFS 
library(rlang)
library(readxl)
library(readr)
library(lubridate)
library(lfstat)
library(tibble)
library(tidyverse)

SMFS_09192023 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_09192023.csv")
SMFS_OrganismsLookUp <- read_excel("Data/SMFS_OrganismsLookUp.xlsx")
SMFS_AgesBySizeMo <- read_excel("Data/SMFS_AgesBySizeMo.xlsx")

str(SMFS_09192023)
SMFS_09192023$Year<-year(SMFS_09192023$SampleDate)
SMFS_09192023$Month<-month(SMFS_09192023$SampleDate)
SMFS_09192023$WaterYear<-water_year(SMFS_09192023$SampleDate, as.POSIX = FALSE, origin="usgs")
```

# Create SMFS_Integrate dataframe
## Filtering
```{r}
#Now I just want otter trawl data
SMFS_OTR_09192023<-SMFS_09192023 %>% filter(MethodCode %in% "OTR") #filters otter trawl samples from table

#Now I want to constrain the data to the years of interest (2011 to now)
str(SMFS_OTR_09192023)
SMFS_OTR_09192023$WaterYear<-as.character(SMFS_OTR_09192023$WaterYear)#when converting the factor to a number, need to first make it a character and THEN numeric. Otherwise, the number gets changedfor some reason.
SMFS_OTR_09192023$WaterYear<-as.numeric(SMFS_OTR_09192023$WaterYear) #need to make water year numeric so I can filter out years easily.
SMFS_OTR_09192023$Year<-as.numeric(SMFS_OTR_09192023$Year) #need to make water year numeric so I can filter out years easily.

SMFS_Integrate<-SMFS_OTR_09192023 %>% filter(Year > 2010, !OrganismCode %in% "NOTRAWL") #Filtering out years which overlap with South Bay Otter Trawl Survey and non-trawls.
```

## Adding species name information

```{r}
SMFS_Integrate<-left_join(SMFS_Integrate, SMFS_OrganismsLookUp, by = "OrganismCode") #Combining the sampling information with organism names

SMFS_Integrate<-mutate(SMFS_Integrate, gensp = paste(Genus, Species, sep=" ")) #making column gensp to match levi. 

#Many rows of NA NA for gensp.
na_gensp_rows <- subset(SMFS_Integrate, gensp == "NA NA")
```

### Unknown Gobies
```{r}
#Would love to resolve these unknown gobies.
unkgobs<-na_gensp_rows %>% filter(OrganismCode %in% "GBU")
GobyIdNeed<-SMFS_Integrate %>% filter(SampleRowID %in% unkgobs$SampleRowID)
#write.csv(GobyIdNeed, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/MissingGobyIDs.csv") #Send to teejay for help?
```

### Fixing gensp for unknown fish
```{r}
unique(na_gensp_rows$CommonName)
na_gensp <-na_gensp_rows %>% group_by(CommonName) %>% summarise(nrows= n(), CountTotal = sum(Count)) #unknown fish IDs. Remove. Losing "Goby Unknown"        "Herring Unknown"     "Unknown Larval Fish" "Heel Splitter"  "Catfish Unknown"     "Water quality only"

#fixing names
SMFS_Integrate$gensp[SMFS_Integrate$OrganismCode=='GBU']<- "Unknown Goby"
SMFS_Integrate$gensp[SMFS_Integrate$OrganismCode=='UKNLARV']<- "Unknown Larval Fish"
SMFS_Integrate$gensp[SMFS_Integrate$OrganismCode=='CFU']<- "Catfish Unknown"
SMFS_Integrate$gensp[SMFS_Integrate$OrganismCode=='CLUPEIDAE']<- "Herring Unknown"
SMFS_Integrate$gensp[SMFS_Integrate$OrganismCode=='HEELSPLITT']<- "Heel Splitter"

unique(SMFS_Integrate$gensp)
#SMFS_Integrate_NoNAFish<-SMFS_Integrate %>% filter(!gensp == "NA NA") # Leaving in for now
```
 CommonName          nrows CountTotal
  <chr>               <int>      <dbl>
1 Catfish Unknown        15         16
2 Goby Unknown          728       7001
3 Heel Splitter           4          4
4 Herring Unknown        28         32
5 Unknown Larval Fish   113        347

## One Measured Fish One Row 
I assume we don't want 1 row, 1 fish for all of the unmeasured fish and would instead prefer 1 row 1 fish for all of the measured fish, and all of the unmeasured fish together in a row to later be expanded once we decide on a method for doing that. Otherwise, I can easily change the code to uncount everything by removing the filter.
```{r}
OneMeasuredFishOneRow <- SMFS_Integrate %>%
  filter(StandardLength > 0) %>% #only keeping measured fish
  uncount(Count, .remove = FALSE) #keeps number of rows that corresponds to the Count value
OneMeasuredFishOneRow$Count<-1 #assigns value of 1 to all counts
unique(OneMeasuredFishOneRow$Class) # Only "Osteichthyes"       "Cephalaspidomorphi"
nrow(OneMeasuredFishOneRow) #71934

#Give these fish IDs
OneMeasuredFishOneRow <- OneMeasuredFishOneRow %>%
  dplyr::mutate(FishID = row_number()) #df becomes size 0. Hopefully that is okay?
```


# Plus Count Expansion
Trawls have a range of measured fish for plus counts. Some of these have a wide range (one 625mm fish and the rest under 200mm for example). 
1. Pull from comments. Or use the comments to ground-truth.
2.Identify and remove outliers from trawls when doing this calculation.
3. Either random assignment or proportional (Levi doing random). Poisson distribution? 
4. Give each fish own row and ID.  

## Separating plus count fish
```{r}
#Bringing plus counts back together with expanded dataframe. 
PlusCountDF<-SMFS_Integrate %>%
  dplyr::filter(!StandardLength > 0)
PlusCountDF <- PlusCountDF %>%
  dplyr::mutate(FishID = NA)
AllSpecimen_MoreRows<-rbind(OneMeasuredFishOneRow, PlusCountDF)
unique(PlusCountDF$Class) #All the classes are back 

#OtherPlus<-PlusCountDF %>% filter(Class %in% c("Malacostraca", "Crustacea", "Bivalvia", "many", "Hydrozoa", "N/A",  "Insecta", "Angiosperms",  "NA"))

#Save New DF
#write.csv(AllSpecimen_MoreRows, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_OneFishOneRow.csv")

PlusTable<-PlusCountDF %>% filter(OrganismCode %in% c("ASH", "SKG", "ISS", "STBK", "LFS", "SCP")) %>% group_by(OrganismCode) %>% dplyr::summarise(CountSum = sum(Count)) #Tells us the amount of plus counts.
print(PlusTable)
```

OrganismCode CountSum
  <chr>           <dbl>
1 ASH                 6
2 ISS               336
3 LFS               696
4 SCP               206
5 SKG               709
6 STBK              380

Built with Levi and Kyle.
## Poisson distribution random sample
```{r}
#Doesn't work well with non-fish. Let's isolate the fish and then run the code. 
PlusCount_OC<-PlusCountDF %>% filter(Class %in% c("Osteichthyes","Cephalaspidomorphi")) #Technically we don't have any plus count Cephalaspidomorphi but in case they appear in the future this code is needed.
#bring back the others later: "Malacostraca" "Crustacea"    "Bivalvia"    [4] "many"         "Hydrozoa"     "N/A"   "Insecta"      "Angiosperms" [10] "NA"    
unique(PlusCount_OC$OrganismCode)
sum(PlusCount_OC$Count)

#ZeroCount<-exp_table %>% filter(Count == 0)
#write.csv(ZeroCount, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_ZeroCount.csv")

exp_table <- 
  OneMeasuredFishOneRow %>% #Only the samples with measured fish
  group_by(SampleRowID,gensp) %>%   # group x taxon & tow 
  dplyr::summarize(m_sl = mean(StandardLength, na.rm=T),  # generate mean sl (for rnorm expansion)
            sd_sl = sd(StandardLength, na.rm=T), # generate sd (for rnorm expansion)
            n_sl = length(StandardLength)) %>%  # generate n of length measurements (as a check)
  full_join(PlusCount_OC, c("SampleRowID", "gensp")) %>%  # join with "Osteichthyes" & "Cephalaspidomorphi" plus counts (number of lengths that require expansion x tow x spp)
  filter(Class %in% c("Osteichthyes","Cephalaspidomorphi"),
         StandardLength == 0, Count != 0) %>% # keep only the plus count data (remove length data). 2 fish have count 0. Sent to Teejay.
  mutate(sd_sl = if_else(Count==1, 0.2*m_sl, sd_sl)) %>%  # if only 1 length value, assume sd = 0.2*SL (if no lengths, then expanded lengths will be NA).
  dplyr::select(SampleRowID, gensp, m_sl, sd_sl, n_sl, StandardLength, Count) %>% 
  data.frame 
# generate unique "sp_id" vector for looping
#exp_table should only have 8 columns: tow id, code, 
exp_table$id_sp <- paste0(exp_table$SampleRowID, "_", exp_table$gensp) 

id_sp <- unique(exp_table$id_sp)
id_sp
# ...expanded lengths list/table ----
exp_table$id_sp
exp_sl_list <- list()   # establish empty expanded lengths list
for(i in id_sp){        # add expanded lengths to list
  # i = id_sp[1]
  temp <- exp_table %>% filter(id_sp ==i)
  set.seed(1)
  df <- data.frame(
    SampleRowID = rep(temp$SampleRowID, temp$Count), #update with column names.
    gensp =  rep(temp$gensp, temp$Count),
    Count = rep(1, temp$Count),
    StandardLength = round(rpois(temp$Count, temp$m_sl)),
    type = "length_plus")
  df_list <- list(df)  # convert to list element
  names(df_list) <- i  # name list element based on id_sp value
  exp_sl_list <- append(exp_sl_list, df_list) # add new element to master list
}
exp_sl_list 

#Okay, now take those lengths and assign them to fish in the SMFS dataset.

#First make this list something I can work with by turning it into a dataframe.

PlusAssigned<-do.call(rbind.data.frame, exp_sl_list) #creates a dataframe from the list (though strange identifier for rows). 

PlusAssigned <-  PlusAssigned %>% 
    rownames_to_column('id_sp')#fixing column name for id_sp
PlusAssigned$PlusID<-1:14983 #I need to track these fish. Giving them each a number for simplicity.
#negative_sl_rows <- PlusAssigned %>% filter(sl<0) #No negative assigned lengths :)

#How do these turn into size bins when I join with the size by month approach. 

#Need to put other columns back in. 
joindf<-PlusCount_OC %>% dplyr::select(-Count, -StandardLength) 
#joindf<-PlusCount_OC %>% dplyr::select(-Count)

PlusJoined<-left_join(PlusAssigned, joindf) %>% distinct() #Okay I've now added month and organism code so I should be able to give ages #14983 fish. 
unique(PlusJoined$CatchComments)

OneFishOneRow<-bind_rows(OneMeasuredFishOneRow, PlusJoined)
#adding age class info. 86917

```
179 fish were unmeasured but had no other measured fish to be based on. They have NA standard lengths.

# Adding Age Class Info
```{r}
#Need to merge with ages. Will move this to data wrangling after plus counts are dealt with.
#Need to update the max in the csv for sizes for HCH to 266 in Jan, BF to 450 in Oct, and ASH to 460 in Aug
SMFS_AgesBySizeMo<-SMFS_AgesBySizeMo %>% mutate(Max = case_when(OrganismCode == "HCH" & Month == 1 ~ 267,
                          OrganismCode == "BF" & Month == 10  ~ 451,
                          OrganismCode == "ASH" & Class == "Age-1+" & Month == 8  ~ 461, TRUE~Max)) #The code above increases the maxes for the 3 fish previously being deleted because the SMFS_AgesBySizeMo does not have ranges that capture them.

#before I add age class, I have to rename a duplicated column name "class"
AgesBySizeMo<-SMFS_AgesBySizeMo %>% rename("AgeClass"="Class")
#ggplot(SMFS_AgesBySizeMo, aes(x=Month, color=Class))+
 # geom_linerange(ymin=SMFS_AgesBySizeMo$Min, ymax=SMFS_AgesBySizeMo$Max)+
 # facet_wrap(~OrganismCode) #I want to double check these are right with no overlap by plotting them but it is hard :(

#If a fish is given no age class I want to assign NA to them
unique(SMFS_AgesBySizeMo$OrganismCode) #33 species
unique(OneFishOneRow$OrganismCode) #52
unique(OneFishOneRow$Class) #"Osteichthyes" & "Cephalaspidomorphi"

#adding age class info
AgedFish_Class<- left_join(OneFishOneRow, AgesBySizeMo, join_by("OrganismCode", "Month")) #210053
unique(AgedFish_Class$OrganismCode) #52

#left join because I don't want age classes for fish we did not catch.

glimpse(AgedFish_Class) #148,341 rows, 52 columns.

# Create a boolean condition based on the size range
condition <- between(AgedFish_Class$StandardLength, AgedFish_Class$Min, AgedFish_Class$Max) #Will not apply to NAs!

# Update the 'Class' column in 'AgedFish_Class'
AgedFish_Class$AgeClass <- ifelse(condition, AgedFish_Class$AgeClass, "Delete") #Keeping age class assignments if condition is met, otherwise assigning "delete" value to the row.
Deleted2<-AgedFish_Class %>% filter(AgeClass %in% "Delete") %>% dplyr::select(StandardLength, Min, Max, AgeClass, OrganismCode, SampleRowID)#123122

unique(AgedFish_Class$AgeClass) # "Age-0"  "Delete" "Age-1+" "Age-1"  "Age-2+" "N/A"  

AgedFish_Class <- filter(AgedFish_Class, AgeClass %in% c("Age-0", "Age-1+", "Age-1",  "Age-2+", "N/A",  NA)) #deleting the rows that did not match the condition
unique(AgedFish_Class$AgeClass) #"Age-0"  "Age-1+" "Age-1"  "Age-2+" "N/A"
AgedFish_Class<-AgedFish_Class %>% distinct()#Not sure why these had some duplicated rows.
```
AgedFish_Class duplicated rows: Plus IDs 14983 (+1), 14982 (+2), 14979 (+1), 14978 (+2), 14977 (+1), 14975 (+1), 14974 (+1), 14971 (+1), 14970 (1), 14969 (+1), 14804 (+2). All had NA standard length. 

```{r}
#Now adding Non-OC counts back in
NonOC<-SMFS_Integrate %>% filter(!Class %in% c("Osteichthyes","Cephalaspidomorphi"))

AgedFish_Class<-AgedFish_Class %>% dplyr::select(-Min, -Max, -...8)

NonOC<-NonOC %>% mutate(FishID= "NotFish", PlusID= "NotFish", type = "NotFish", AgeClass = NA)

NonOC$id_sp <- paste0(NonOC$SampleRowID, "_", NonOC$gensp)
colnames(AgedFish_Class)
colnames(NonOC)

AllCatch_AgeClasses<-rbind(AgedFish_Class, NonOC) #107307
AllCatch_AgeClasses<-AllCatch_AgeClasses %>% distinct()
sum(AllCatch_AgeClasses$Count) #731505
sum(SMFS_Integrate$Count)#731505
```

#### Comparing and fixing plus assignments based on Catch comments
```{r}
#Compare to comments. How often is yoy correctly assigned?
#last Catch Comments...

#Which samples_ids have comments providing info on sizes?
unique(AllCatch_AgeClasses$CatchComments)
AllYOY<-AllCatch_AgeClasses %>% filter(CatchComments%in%c("YOY <40 mm SL", "YOY", "young of year", "yoy","all YOY (<30  mm)","all YOY (<40 mm SL)","all YOY (<40 mmSL)", "all YOY (<35 mm)", "YOY-sized fish (~45 mm SL)", "age 0","all young of year", "YOY (<135 mm SL)", "YOY (<40 mm SL)","YOY (<50 mm SL)", "age 0 (<40 mmm)", "YOY (<30 mm)", "(YOY - all less than 100 mm SL)", "age-o fish (<30 mm SL)", "young of year;young of year", "YOY (<60 mm SL)", "age-0 fish","YOY (<50 mmSL)", "all YOY (<100 mm SL)", "all YOY")) #Only fix the young of year. 
#Create for_loop that filters for each of these samples and create a new column with adjusted standard lengths and bins...
YoyList<-AllYOY$id_sp #69 id_sp combos that are all YOY plus counts
YoyList
#Okay, now I have a list of id_sp that should correlate.
Comparison<-AllCatch_AgeClasses$id_sp
Comparison

ShouldYOY<-AllCatch_AgeClasses %>% filter(id_sp %in% YoyList)
#Save list of incorrect fish.
ShouldYOYButNotYOY<-ShouldYOY %>% filter(!AgeClass%in%"Age-0")
ShouldYOYButNotYOY$PlusID

#Make the not yoy but should yoy turn into yoy since we know they should be...
AllCatch_AgeClasses$AgeClass[AllCatch_AgeClasses$PlusID %in% ShouldYOYButNotYOY$PlusID]<-"Age-0" #fixing age class
AllCatch_AgeClasses$StandardLength[AllCatch_AgeClasses$PlusID %in% ShouldYOYButNotYOY$PlusID]<-40 #fixing sl. 40 should keep all of these fish squarely in the yoy category. 
ShouldYOY2<-AllCatch_AgeClasses %>% filter(id_sp %in% YoyList)
ShouldYOYButNotYOY2<-ShouldYOY2 %>% filter(!AgeClass%in%"Age-0")#Done

#There are some fish with no length assignment because they have no measured fish to be based off of. They were one-offs that length measurements were missed on. Levi wants me to deal with them by basing their lengths on lengths of the day.

```

For presence/absence model, filter on tows that measured big fish. Remove those fish/observations. And then for all tows, label each tow as 'presence' or 'absent'. If there was a tow that only measured big fish, it will look like no fish were captured because you removed all instances of big fish measurements. This avoids dealing with plus count, which we don't need for the presence/absent model. I



#### How does possion compare to original distribution? (INCOMPLETE CODE)
```{r}
#Trying to see how the poisson distribution compares to the original distribution of lengths for the same samples/species.
id_sp2 <- unique(SMFS_Integrate_NoNAFish_Expand$id_sp)

tSample<-sample(id_sp2, 30, replace=FALSE)

PlusAssigned_random <-PlusAssigned %>% filter(id_sp %in% tSample)

SMFS_Integrate_NoNAFish_Expand %>%
  group_by(SampleRowID,gensp) %>% ggplot(aes(x=StandardLength))+geom_density()
SMFS_Integrate_NoNAFish_Expand$id_sp <- paste0(SMFS_Integrate_NoNAFish_Expand$SampleRowID, "_", SMFS_Integrate_NoNAFish_Expand$gensp) 
glimpse(SMFS_Integrate_NoNAFish_Expand)
#Randomly sample these id's because doing all of them takes too long...
SMFS_Integrate_NoNAFish_Expand_random<- SMFS_Integrate_NoNAFish_Expand%>% filter(id_sp %in% tSample)

for(i in tSample){
  # Filter your data based on the current ID
  subset_data <- SMFS_Integrate_NoNAFish_Expand %>%
    filter(id_sp == i)
  
  p<-ggplot(subset_data, aes(x=StandardLength))+geom_histogram(binwidth = 1)+labs(title = paste("ID:", i))
  
  print(p)
}

  
p<-SMFS_Integrate_NoNAFish_Expand_random %>% group_by(id_sp) %>% summarize(MEAN=mean(StandardLength), SD = sd(StandardLength))
p
```

#### Final Check
I should have the same number of fish that I started with...
```{r}
glimpse(AllCatch_AgeClasses) #55 columns. 107,307 observations
glimpse(SMFS_Integrate) #49 columns. 80,044 observations
sum(AllCatch_AgeClasses$Count) # 731505
sum(SMFS_Integrate$Count)#731505
```


# Adding zeros - LAST STEP
```{r}
#Now add zeros!

#I need to make a dataframe that repeats the age classes for each organism for each sampling event. 

#First, clean up the agesbysizemo df to only contain needed information. 

Ages<-SMFS_AgesBySizeMo %>% dplyr::select(Class, OrganismCode) %>% rename(AgeClass = Class) %>% filter(AgeClass %in% c("Age-0",  "Age-1+", "Age-1",  "Age-2+")) %>%   distinct()

#Now for every sampling event, repeat all of the species x age combinations. Make a df of the sampling events.
SamplingEvents<-SMFS_Integrate %>% dplyr::select(SampleRowID, StationCode, SampleDate, SampleTime, WaterTemperature, Salinity, DO, PctSaturation, Secchi, SpecificConductance, TideCode, UserName, ElecCond, TrawlRowID, TowDuration, TrawlComments, Year, Month, WaterYear) %>% distinct()

# Create all combinations of sampling events with ages
AllCombinations <- crossing(SamplingEvents, Ages)

unique(AllCatch_AgeClasses$AgeClass)
# Separate rows with OrganismCode that lack AgeClass information
has_ageclass_info <- AllCatch_AgeClasses %>% filter(AgeClass %in% c("Age-0",  "Age-1+", "Age-1",  "Age-2+"))

no_ageclass_info <- AllCatch_AgeClasses %>% filter(AgeClass %in% c("N/A",  NA))

NewCombinations <- AllCombinations %>%
  anti_join(has_ageclass_info, by = c("SampleRowID", "StationCode", "SampleDate", "SampleTime", "WaterTemperature", "Salinity", "DO", "PctSaturation", "Secchi", "SpecificConductance", "TideCode", "UserName", "ElecCond", "TrawlRowID", "TowDuration", "TrawlComments", "Year", "Month", "WaterYear", "OrganismCode", "AgeClass")) %>% mutate(Count=0, type="Absence")


# Perform join and complete missing AgeClass for rows with AgeClass information
completed_with_ageclass <- bind_rows(NewCombinations, has_ageclass_info)

#Do the same for the non-aged specimens so we get 0's for those when absent as well. 
no_ageclass_info <- AllCatch_AgeClasses %>% filter(AgeClass %in% c("N/A",  NA)) 

#Just want the organism codes
NonAgedOrganisms<-unique(no_ageclass_info$OrganismCode)

NonagedCombinations <- crossing(SamplingEvents, NonAgedOrganisms) 
NonagedCombinations$OrganismCode <-NonagedCombinations$NonAgedOrganisms
NewnewCombinations <- NonagedCombinations %>%
  anti_join(no_ageclass_info, by = c("SampleRowID", "StationCode", "SampleDate", "SampleTime", "WaterTemperature", "Salinity", "DO", "PctSaturation", "Secchi", "SpecificConductance", "TideCode", "UserName", "ElecCond", "TrawlRowID", "TowDuration", "TrawlComments", "Year", "Month", "WaterYear", "OrganismCode")) %>% mutate(Count=0, type="Absence", AgeClass=NA)
completed_NA_ageclass <- bind_rows(NewnewCombinations, no_ageclass_info)

Massive<-bind_rows(completed_with_ageclass, completed_NA_ageclass)
Massive_Codes<-Massive %>% dplyr::select(!c("Native", "Resident", "CommonName", "Phylum", "Class", "Order", "Family", "Genus", "Species", "Comments"))

Massive_Codes<-left_join(Massive_Codes, SMFS_OrganismsLookUp, by="OrganismCode")
Massive_Codes<-mutate(Massive_Codes, gensp = paste(Genus, Species, sep=" "), id_sp = paste(SampleRowID, gensp, sep="_")) %>% dplyr::select(!c("NonAgedOrganisms", "OrganismRowID.y", "OrganismRowID.x"))
unique(Massive_Codes$OrganismCode)
sum(Massive_Codes$Count) #731505
```

# Save
```{r}
#Save New DF
write.csv(Massive_Codes, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_WithAbsences_062024.csv")
```


# Investigating NA's in dataset (DOES NOT REMOVE THEM)
```{r}
#How many rows will we lose if we remove NAs?
sum(is.na(SMFS_Integrate$WaterTemperature)) #44
sum(is.na(SMFS_Integrate$DO)) #495
sum(is.na(SMFS_Integrate$Salinity)) #47
sum(is.na(SMFS_Integrate$Secchi)) #100
#This isn't super helpful... We have multiple rows for one 
Na_Assessment<-SMFS_Integrate %>% 
  group_by(SampleRowID) %>% 
  dplyr::summarize(WaterTemp_NA=sum(is.na(WaterTemperature)),
            DO_NA =  sum(is.na(DO)),
            Sal_NA = sum(is.na(Salinity)),
            Secchi_NA=sum(is.na(Secchi)))
Na_Assessment<- Na_Assessment %>% filter(WaterTemp_NA>0 | DO_NA>0 | Sal_NA>0 | Secchi_NA>0)
#Overall, there are 46 sampling events with missing wq data.
#let's find out more about these.
NaSamples<-Na_Assessment$SampleRowID

SMFS_OTR_Thesis_WQNA<-SMFS_Integrate %>% filter(SampleRowID %in% NaSamples) %>% group_by(SampleRowID) %>% dplyr::select(SampleRowID, StationCode, MethodCode, TowDuration, WaterTemperature, Secchi, Salinity, DO, PctSaturation, SpecificConductance, TideCode, SampleDate, Year, WaterYear)
SMFS_OTR_Thesis_WQNA<-as.data.frame(SMFS_OTR_Thesis_WQNA)
SMFS_OTR_Thesis_WQNA<-SMFS_OTR_Thesis_WQNA %>% distinct()
#Save New DF
#write.csv(SMFS_OTR_Thesis_WQNA, "C:/Users/cnewe/OneDrive/Documents/Incubator/Code/FishForesightIncubator/Data/SMFS_NA_WaterQuality.csv") #This holds the information with samples that have NA wq data


#Visualizing fish counts by year for each species
FishCountTable<-SMFS_Integrate %>% group_by(OrganismCode) %>% summarise(Total=sum(Count))

FishCountTable<-SMFS_Integrate %>% group_by(OrganismCode, Year) %>% summarise(Total=sum(Count))

FishCountTable %>% spread(key = Year, value = Total)

#Adding zeros for sampling events with no fish caught
#https://derekogle.com/fishR/2018-04-19-Adding-Zero-Catches
length(unique(SMFS_Integrate$SampleRowID)) 
length(unique(SMFS_Integrate$TrawlRowID)) 
length(unique(SMFS_Integrate$OrganismCode))
3702*79


Nas_<-SMFS_Integrate %>% filter(TowDuration %in% NA)
NA_SampleRowIDs<-unique(Nas_$SampleRowID)
NA_Tow_Info<-SMFS_09192023 %>% filter(SampleRowID %in% NA_SampleRowIDs)
unique(NA_Tow_Info$SampleRowID)
#Okay I am deducing that these samples are actually fake samples with incidental or rod and reel info? 
#Removing all NA Trawl Duration (below)


#Why NA DO?
NoDO<-SMFS_Integrate %>% filter(DO %in% NA)
#Reasons for missing DO: 
#Row 81, 232, 253,  - just didn't record DO or PctSat
#Is it missing for other organisms from the same trawl?

NoDOTrawlID<-unique(NoDO$TrawlRowID)
length(NoDOTrawlID) #35 trawls
NoDOTrawls<-SMFS_Integrate %>% filter(TrawlRowID %in% NoDOTrawlID) #nope. The trawl just doesn't have the info! Remove these rows.


#Now checking salinity
NoSal<-SMFS_Integrate %>% filter(Salinity %in% NA)
NoSalTrawlID<-unique(NoSal$TrawlRowID)
length(NoSalTrawlID) #6 trawls
NoSalTrawls<-SMFS_Integrate %>% filter(TrawlRowID %in% NoSalTrawlID) #These trawls just doesn't have the info! Remove these rows.

#Now checking Secchi
NoSecchi<-SMFS_Integrate %>% filter(Secchi %in% NA)
NoSecchiTrawlID<-unique(NoSecchi$TrawlRowID) 
length(NoSecchiTrawlID) #3 trawls
NoSecchiTrawls<-SMFS_Integrate %>% filter(TrawlRowID %in% NoSecchiTrawlID) #These trawls just doesn't have the info! Remove these rows.

#Now checking Water Temp
NoWT<-SMFS_Integrate %>% filter(WaterTemperature %in% NA)
NoWTTrawlID<-unique(NoWT$TrawlRowID) 
length(NoWTTrawlID) #5 trawls
NoWTTrawls<-SMFS_Integrate %>% filter(TrawlRowID %in% NoWTTrawlID) #These trawls just doesn't have the info! Also missing all other WQ. Remove these rows.
#I remove these NAs in the SMFS_WranglingForThesis rmd
```
